---
output:
    html_document:
            keep_md: yes
            toc: true
            toc_float: true
            code_folding: hide
---


## Note


- R has to be run in interactive session

- The default sinteractive allocation is 1 core (2 CPUs) and 4GB of memory.
  This script requires more memory. Set `--mem=16g` in order to avoid
  segmentation fault error which occurs during importing large dataset (e.g.
  HDF5 file).

## Packages


```{r global_options, include=FALSE}
knitr::opts_chunk$set(
    warning=FALSE,
    message=FALSE,
    results='hide'
    )
```


```{r load_packages}
library(readr)
library(knitr)
library(Signac)
library(Seurat)
library(ensembldb)
library(EnsDb.Hsapiens.v86)
library(BSgenome.Hsapiens.UCSC.hg38)
library(GenomicRanges)
library(biovizBase)
library(future)

set.seed(127)
```


```{r parallelize, include=FALSE}

# Assign the number of cores for parallelization
ncpus <- future::availableCores()

# Adjust max memory allowed
options(future.globals.maxSize=40000 * 1024^2,
        mc.cores=ncpus)

# Parallelize the run
plan("multicore", workers=ncpus)


```


## Build annotation using EnsDb and GetGRangesFromEnsDb()


```{r annotate_genes_ensdb, results='asis'}

# Change the seqlevels to USCS style
seqlevelsStyle(EnsDb.Hsapiens.v86) <- "UCSC"

# Annotate genes using GetGRangesFromEnsDb() from Signac and EnsDb
# This function requires the biovizbase package installed (conda: bioconductor-biovizbase)
annotation.ensdb <- GetGRangesFromEnsDb(ensdb=EnsDb.Hsapiens.v86)

genome(annotation.ensdb) <- "hg38"

```

### EnsDb annotation summary


```{r ensdb_summary, results='asis'}

# Explore the annotation object
# Note: make sure you have gene_biotype column.
cat(summary(annotation.ensdb))

```

### EnsDb annotation exploration


```{r ensdb_print, results='asis'}

kable(head(data.frame(annotation.ensdb)))


```


```{r save_annotation_output}

write_rds(annotation.ensdb, "annotation.rds")

```


## Session Info

```{r session_info}

sessionInfo()
```


