---
output:
    html_document:
            keep_md: yes
            toc: true
            toc_float: true
            code_folding: hide
title: "10X Genomics Multiome - Test Dataset Demonstration"
---


## Note

- This demonstration is designed to validate a test dataset analyzing small groups of cells, genes, and peaks from 10X multiome

- Full dataset: [PBMC from a healthy donor - granulocytes removed through cell sorting (3k)](https://www.10xgenomics.com/resources/datasets?query=&page=1&configure%5Bfacets%5D%5B0%5D=chemistryVersionAndThroughput&configure%5Bfacets%5D%5B1%5D=pipeline.version&configure%5BhitsPerPage%5D=500&menu%5Bproducts.name%5D=Single%20Cell%20Multiome%20ATAC%20%2B%20Gene%20Expression) provided by 10X Genomics

- Workflow reference: [Joint RNA and ATAC analysis: 10x multiomic](https://satijalab.org/signac/articles/pbmc_multiomic.html) demonstrated by Satija Lab


```{r global_options, include=FALSE}
knitr::opts_chunk$set(
    warning=FALSE,
    message=FALSE,
    results='hide'

    )
```


## Loading packages

- [Signac (v1.3.0)](https://satijalab.org/signac/)

- [Seurat (v4.0.3)](https://satijalab.org/seurat/)


```{r load_packages}
library(knitr)
library(tidyverse)
library(Signac)
library(Seurat)
library(EnsDb.Hsapiens.v86)
library(BSgenome.Hsapiens.UCSC.hg38)
library(plotly)
library(future)
library(ggrepel)
library(patchwork)

set.seed(127)
```


## Set parallelization

- Parallelization in this workflow uses the [future](https://cran.r-project.org/web/packages/future/index.html) package. 

- Reference: [Parallelization in Seurat with future](https://satijalab.org/seurat/articles/future_vignette.html)



```{r parallelize, include=FALSE}

# Assign the number of cores for parallelization
ncpus <- future::availableCores()

# Adjust max memory allowed
options(future.globals.maxSize=48000 * 1024^2,
        mc.cores=ncpus)

# Parallelize the run
plan("multicore", workers=ncpus)


```

## Import path variables


```{r path_variables}

source("path.R")

```

## Import the GEX and ATAC matrices

```{r import_counts}

# Import count data
counts <- readRDS(to.matrix)

# Explore the output
class(counts)  # Data type is sparse matrix saving memory
summary(counts)
```

## Create and import annotation

```{r import_annotation}


# Import the annotation object
annotation <- readRDS(to.annotation)


```

## Create a Seurat object and add the ATAC assay {.tabset .tabset-pills}


```{r create_seurat}

# Create a Seurat object
seurat <- CreateSeuratObject(counts=counts$`Gene Expression`,
                             assay="RNA")

# Add mitochondrial read to the seurat object
seurat[["percent.mt"]] <- PercentageFeatureSet(seurat,
                                               pattern="^MT-")

# Create ATAC assay and add it to the object
#
# Note:
# The CreateChromatinAssay() function filters out cells with 0 (pooled) atac counts
# (e.g. counts <- counts[, colSums(counts) > min.features] where mean.features is set to 0 by default])
# whereas the CreateSeuratObject() function keeps all the cells by shifting the min.features argument to -0.1
# So this arguments in the CreateChromatinAssay() function has to be manually set to -0.1
seurat[["ATAC"]] <- CreateChromatinAssay(counts=counts$Peaks,
                                         sep=c(":", "-"),
                                         fragments=fragpath,
                                         annotation=annotation,
                                         min.features = -0.1) # IMPORTANT!





# Explore the object
seurat
# e.g. Expected output is:
# 134920 features across 2711 samples within 2 assays
# Active assay: RNA (36601 features, 0 variable features)
# 1 other assay present: ATAC



```

### Seurat metadata

```{r explore_seurat_head, results='asis'}

# Message the number of cells in the dataset
cat(paste("Total", nrow(seurat@meta.data), "cells are analyzed as shown below"))

# Explore metadata
kable(head(seurat@meta.data))

```

### Seurat GEX counts


```{r explore_seurat_gex_counts, results='asis'}

# Message the number of genes in the dataset
cat(paste("Total", nrow(counts[[1]]), "genes are analyzed as shown below"))

# Explore the GEX assay data
kable(GetAssayData(seurat[["RNA"]], slot="counts")[1:10, 1:5])

```

### Seurat ATAC counts

```{r explore_seurat_atac_counts, results='asis'}

# Message the number of genes in the dataset
cat(paste("Total", nrow(counts[[2]]), "peaks are analyzed as shown below"))


# Explore the ATAC assay data
kable(GetAssayData(seurat[["ATAC"]], slot="counts")[1:10, 1:5])

```

## Quality control {.tabset .tabset-pills}



```{r quality_control}

# Set the default assay to ATAC
DefaultAssay(seurat) <- "ATAC"

# Explore the output
seurat
# e.g. Expected output is:
# An object of class Seurat
# 123703 features across 5425 samples within 2 assays
# Active assay: ATAC (87102 features, 0 variable features)
# 1 other assay present: RNA


# Calculate the strength of the nucleosome signal per cell
# to fragments < 147bp (nucleosome-free) are calculated
seurat <- NucleosomeSignal(seurat)

# Explore the calculated nuleosome signal in the first 20 cells
# (returns a numeric vector)
head(seurat$nucleosome_signal, 20)

# Compute the transcription start site (TSS) enrichment score for each cell
# The computed score will be added to the object metadata as "TSS.enrichment"
seurat <- TSSEnrichment(seurat)

# Explore the new column storing TSS enrichment score
head(seurat$TSS.enrichment)

# Imput NA values
# (This step is not needed for full datasets)
sum(is.na(seurat@meta.data)) # the number of NAs
seurat@meta.data[is.na(seurat@meta.data)] <- 0  # Imput


```

### QC metrics before filtering

```{r before_qc_filtering, results='asis', fig.width=12, fig.height=12}


# Plot the QC metrics before filtering
VlnPlot(object=seurat,
        features=c("nCount_RNA",
                   "percent.mt",
                   "nCount_ATAC",
                   "TSS.enrichment",
                   "nucleosome_signal"),
        ncol=4,
        pt.size=0.5)



```


## Session Info

```{r session_info, class.output="pre"}

# Save the seurat object
write_rds(seurat, "seurat_qc.rds")

sessionInfo()

```


