---
title: "Peakcalling - II"
output:
    html_document:
            toc: true
            toc_float: true
            code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    warning=FALSE,
    message=FALSE
    )

```

```{r packages_and_functions, include=FALSE}
library(Signac)
library(Seurat)
library(GenomeInfoDb)
library(knitr)
library(future)
library(GenomicRanges)
set.seed(1234)

source('common.R')
```

## Parallelization

```{r parallelize, include=FALSE}

# Assign the number of cores for parallelization
ncpus <- future::availableCores()

# Adjust max memory allowed
options(future.globals.maxSize=40000 * 1024^2,
        mc.cores=ncpus)

# Parallelize the run
plan("multicore", workers=ncpus)


```

## Set variables

We'll get all the sample and replicate specific variables listed in samples.tsv and units.tsv, respectively. We'll also set paths to input files.

```{r variables}

anno_path <- params[['annotation']]
genome <- params[['genome']]
fragpath <- params[['fragments']]
peaks <- params[['peak']]   # macs2 peak (.narrowPeak)
```


## Read in data

Read seurat and annotation

```{r input}
seurat <- readRDS(params[["seurat"]])

print('Input seurat:')
print(seurat)

annotation <- readRDS(anno_path)
```


## Importing MACS2-called peaks

Blacklist datasets are provided by Signac package. For more info, see 
[this](https://satijalab.org/signac/reference/index.html)

Currently adds MACS peaks to an assay called MACS. Probably would make workflow more interpretable if user 
chooses only one method, ZINBA or MACS, in config. If this is the case, overwrite Peaks assay with MACS 
counts (see comments below).

```{r macs}
# set effective genome size for MACS2
# and
# Remove peaks on nonstandard chromosomes and in genomic blacklist regions
# blacklist_hg38_unified: blacklist dataset included in Signac package
# More blacklist datasets: blacklist_mm10, blacklist_ce10, blacklist_ce11,
# blacklist_dm3, blacklist_dm6, blacklist_hg19, blacklist_hg38
if (genome == 'mm10') {
    effective_size <- 1.87+09
    blacklist <- blacklist_mm10
} else if (genome == 'hg38') {
    effective_size <- 2.7e+09
    blacklist <- blacklist_hg38_unified
}

# Import MACS2-called peaks
peak.df <- read.table(peaks)

# Convert the loci to GRanges obj
peaks <- GRanges(peak.df[,1], IRanges(peak.df[,2], peak.df[,3]))

# Preprocess the peaks
peaks <- keepStandardChromosomes(peaks, pruning.mode = "coarse")
peaks <- subsetByOverlaps(x = peaks, ranges = blacklist, invert = TRUE)

print(paste("Number of peaks after filtering:", length(peaks)))

# Quantify counts in each peak
peak_counts <- FeatureMatrix(fragments=Fragments(seurat),
                              features=peaks,
                              cells=colnames(seurat))


# seurat object may contain a Fragments object
if (is.null(Fragments(seurat))) {
# Add a new assay with the called peak to the seurat object
    seurat[["MACS"]] <- CreateChromatinAssay(counts=peak_counts,
                                             fragments=fragpath,
                                             annotation=annotation)
} else {
    # seurat object may contain cells/barcodes that do not have peaks
    # split object, add macs peaks, combine
    cells_use <- lapply(Fragments(seurat), function(x) {
        names(x@cells)
    })
    cells_use <- unique(Reduce("c", cells_use))
    
    # seurat object has same number of cells as MACS peak count matrix?
    if (length(cells_use) == length(colnames(seurat))) {
        seurat[["MACS"]] <- CreateChromatinAssay(
            counts=peak_counts, 
            fragments=Fragments(seurat),
            annotation=annotation)
    } else {

        # Subset the seurat obj by presence of peaks
        seurat_peaks <- subset(x = seurat, cells = cells_use)
        seurat_no_peaks <- subset(x = seurat, cells = cells_use, invert = TRUE)

        # Add MACS2-called peaks to the obj with peaks
        seurat_peaks[["MACS"]] <- CreateChromatinAssay(
            counts=peak_counts, 
            fragments=Fragments(seurat_peaks),
            annotation=annotation)
        
        # Prep a pseudo matrix for the other obj without peaks
        pseudo_matrix <- matrix(
            0,
            nrow=nrow(seurat_no_peaks),
            ncol=ncol(seurat_no_peaks)
        )

        # Add row and column names
        rownames(pseudo_matrix) <- rownames(seurat_no_peaks)
        colnames(pseudo_matrix) <- colnames(seurat_no_peaks)

        # Add the pseudo matrix to "MACS" assay in the seurat obj without peaks
        seurat_no_peaks[['MACS']] <- CreateAssayObject(counts=pseudo_matrix)

        # Merge both Seurat obj
        seurat <- merge(seurat_no_peaks, seurat_peaks)
    }
}
```

## Save Seurat object

```{r save}
print('Updated Seurat object:')
print(seurat)

print(paste('Updated Seurat object saved in', params[['seurat_out']]))
saveRDS(object=seurat, file=params[['seurat_out']])
```


## Session Info

```{r session_info}
sessionInfo()
```



