---
title: "Building Annotation from User Provided GTF"
output:
    html_document:
            keep_md: yes
            toc: true
            toc_float: true
            code_folding: hide
---


## Note

- R has to be run in interactive session

- The default sinteractive allocation is 1 core (2 CPUs) and 4GB of memory.
  This script requires more memory. Set `--mem=16g` in order to avoid
  segmentation fault error which occurs during importing large dataset (e.g.
  HDF5 file).




```{r global_options, include=FALSE}
knitr::opts_chunk$set(
    warning=FALSE,
    message=FALSE,
    results='hide'
    )
```

## Packages


```{r load_packages}
library(readr)
library(knitr)
library(Signac)
library(Seurat)
# library(ensembldb)
# library(EnsDb.Hsapiens.v86)
# library(BSgenome.Hsapiens.UCSC.hg38)
library(GenomicRanges)
# library(biovizBase)
library(future)

set.seed(127)
```


```{r parallelize, include=FALSE}

# Assign the number of cores for parallelization
ncpus <- future::availableCores()

# Adjust max memory allowed
options(future.globals.maxSize=40000 * 1024^2,
        mc.cores=ncpus)

# Parallelize the run
plan("multicore", workers=ncpus)


```


## Loading path variables

```{r rule_variables}

print('yaml params:')
print(params)

annotation.file <- params[['anno_file']]
genome <- params[['genome']]


```



## Build annotation from generic GTF


```{r annotation_from_generic_GTF, cache=TRUE}
# use cellranger supplied genome to determine whether TSSEnrichment takes any generic annotation file
# as long as in GRanges format and seqlevels match feature coordinates

gtf.ranges <- rtracklayer::import(con = annotation.file)

# common filtering steps for EnsDb and standard GTF objects:
# add generic "gene_biotype" column, Ensdb.Species uses gene_biotype other repos (i.e. Ensembl) use "gene_type"
# Also avoids the following:
# Error in GetTSSPositions(ranges = annotations) : Gene annotation does not contain gene_biotype information
if ( any(colnames(mcols(gtf.ranges)) %in% "gene_type") ) {
  gtf.ranges$gene_biotype <- gtf.ranges$gene_type
}

```


### Pseudogene types

```{r annotation_from_GTF_pseudogenes}

# Optional
# Use all gene/transcript types except pseudogenes:
pseudogene.types <- unique(gtf.ranges$gene_biotype)[grep(".*pseudogene.*", unique(gtf.ranges$gene_biotype) )]

print("Pseudogene types:")
for (g in pseudogene.types) {
    print(g)
}

# Remove unwanted genes
remove.genes <- c("TEC", pseudogene.types, "Mt_.*")
remove.index <- grepl(paste(remove.genes, collapse="|"), gtf.ranges$gene_biotype)
gtf.ranges <- gtf.ranges[ !remove.index ]

# seqlevelsStyle(gtf.ranges) <- 'UCSC'

# Ootional
# remove non-standard chromosomes and mitochondrial (has no chromatin)
# Does not affect TSSEnrichment() but should be done for downstream analyses.


if (genome == 'mm10') {
    genomeStyles.species = 'Mus_musculus'
} else if (genome == 'hg38') {
    genomeStyles.species = 'Homo_sapiens'
}



# genomeStyles.species = 'Mus musculus'
gtf.ranges <- keepStandardChromosomes(
    gtf.ranges,
    species = genomeStyles.species,
    pruning.mode = 'coarse')
gtf.ranges <- gtf.ranges[!seqnames(gtf.ranges) == "chrM"]


```


### Seqlavels


```{r annotation_from_GTF_seqlevels, results='asis'}

cat("seqlevels for working gtf.ranges anno:")
cat(seqlevels(gtf.ranges))
cat(unique(seqnames(gtf.ranges)))


# # lets explore the contigs of reference and Peaks countsL
# # reference:
# gtf_contigs = unique(as.character(seqnames(gtf.ranges)))

# # unique contigs present in Multiome Peaks matrix:
# data_contigs = unique(unlist(strsplit(x = rownames(counts$`Peaks`), split = ':'))[c(TRUE, FALSE)])

# # all contigs in Peaks matrix are present in 10X gene annotation
# all(gtf_contigs %in% data_contigs)

gtf.genes = gtf.ranges[gtf.ranges$type == "gene"]
annotation <- gtf.genes

# genome(annotation) <- "hg38"
genome(annotation) <- genome

```

###  Annotation summary


```{r annotation_from_GTF_summary, results='asis'}
# Explore the annotation object
# Note: make sure you have gene_biotype column.
cat(summary(annotation))

```

### Annotation exploration


```{r annotation_from_GTF_print, results='asis'}

kable(head(data.frame(annotation)))

```


```{r save_annotation_output}
# write_rds(annotation, "annotation.rds")
write_rds(annotation, params[['anno_atac']])
```


## Session Info

```{r session_info}

sessionInfo()
```

