---
title: "chooseR"
output:
    html_document:
        toc_float: true
        toc: true
        code_folding: hide


---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=FALSE,
                      message=FALSE)
```

## Activate conda env

Ensure getting all the packages in `requirements.txt` installed and running `activate conda ./env`

```{r packages_and_functions, include=FALSE}
library(tidyverse)
library(data.table)
library(ggplot2)
library(Signac)
library(Seurat)
library(knitr)
library(Matrix)
library(cluster)
library(purrr)
# library(future)
# library(parallel)

set.seed(1327)  # added to remove unwanted randomness

# Set correct paths
source("chooser_var.R")
```





## Set variables

We'll get all the sample and replicate specific variables listed in samples.tsv and units.tsv, respectively. We'll also set paths to input files.



```{r variables}
config_path = params[['config']]
res <- as.numeric(params[["res"]])
```


## Load Seurat object

- **Input data** = an RDS file storing a **seurat object** (after SCTransform & PCA)


```{r load_data}
config = yaml::read_yaml(file = config_path)

# Read a QC-ed seuat object
obj <- readRDS(params[["seurat_in"]])


# Explore the output
obj

```


## Set variables

```{r set_variables}

# Define the number of PCs to use, and which assay and reduction to use.
# We recommend testing a broad range of resolutions
# For more on picking the correct number of PCs, see:
# https://satijalab.org/seurat/v3.1/pbmc3k_tutorial.html
# npcs <- obj@commands$RunPCA.SCT@params$npcs
# assay <- "SCT"
# reduction <- "pca"


# target = basename(params[['seurat_out']])
# target = gsub("^(.*).clustered_data.rds", "\\1", target)
target <- params[["seurat_out"]]   # changed by Mira
## #  print("Choosing for group:")
cat(target)

groups = config[['chooser']][['groups']]

idx = which(names(groups) %in% target)
 
if (length(idx) == 0) {
    cat("Group is not present in config's normalize groups list.")
    q()
} else if (length(idx) > 1) {
    cat("You are trying to normalize group/assay multiple times. Please fix.")
    q()
}

group_name = names(groups)[idx]
cat('Current group:')
cat(group_name)

group = groups[[idx]]

# determined by ElbowPlots from normalize/integrate rules
npcs = group[['npcs']]

cat('npcs')
cat(npcs)

targets = c("pca", "lsi")
idx = grep(paste(targets,collapse="|"), names(obj@reductions))
reduction = names(obj@reductions)[idx]

# npcs = ncol(obj[[reduction]]@cell.embeddings)
assay = obj[[reduction]]@assay.used

# Create a directory for output plots
# create sample subdir - May be multiple chooseR scripts running at once.
results_path = params[['results_path']]
cat(results_path)

system(paste0("mkdir -p ./", results_path))

```

## Run chooseR

- **Input = seurat object (after SCTransform and PCA)**


```{r chooser}


# Run pipeline
message(paste0("Clustering ", res, "..."))
message("\tFinding ground truth...")

# "Truths" will be stored at glue::glue("{reduction}.{assay}_res.{res}")
obj <- find_clusters(obj,
                   npcs=npcs,
                   reduction = reduction,
                   assay = assay,
                   resolution = res)


# Now perform iterative, sub-sampled clusters (returns a data frame)
results <- multiple_cluster(obj,
                          n = nIter,
                          npcs = npcs,
                          res = res,
                          reduction = reduction,
                          assay = assay)



# Explore the clusters (first 5 rows and columns)
kable(results[1:5, 1:5])
dim(results)



# Extract sampled barcodes
sampled.cells <- results$cell

# Explore the number of sampled cells
length(sampled.cells)

# Now calculate the co-clustering frequencies
message(paste0("Tallying ", res, "..."))

# This is the more time efficient vectorisation
# However, it exhausts vector memory for (nearly) all datasets
# matches <- purrr::map(columns, find_matches, df = results)
# matches <- purrr::reduce(matches, `+`)

# CR change:
results = results[,-1]
columns <- colnames(results)

mtchs <- matrix(0, nrow = dim(results)[1], ncol = dim(results)[1])

i <- 1 # Counter

for (col in columns) {
message(paste0("\tRound ", i, "..."))
mtchs <- Reduce("+", list(
  mtchs,
  find_matches(col, df = results)
))
i <- i + 1
}

message(paste0("Scoring ", res, "..."))
# Original code: # returns class "tibble", all cols type "double"
# matches <- dplyr::mutate_all(
# dplyr::as_tibble(matches),
# function(x) dplyr::if_else(Re(x) > 0, percent_match(x), 0)
# )
# Converted to (by Mira): # returns class "matrix" of type "double"
mtchs <- apply(as.data.table(mtchs),
               2,
               function(x) dplyr::if_else(Re(x) > 0, percent_match(x), 0))

# Sometimes next line causes error for some reason
# mtchs[1:5, 1:5]
print("Scores:")
print(dim(mtchs))


# Now calculate silhouette scores
message(paste0("Silhouette ", res, "..."))


# Subset the seurat obj by sampled cells
if (length(sampled.cells) > 0) {
    sampled.obj <- obj[, sampled.cells]
} else {
    sampled.obj <- obj
}



# Extract clusters from the sampled seurat
clusters <- unlist(sampled.obj[[glue::glue("{reduction}.{assay}_res.{res}")]])

# Compute silhouette width
# NOTE: cluster::silhouette() requires
# either dist (distance matrix) or dmatrix (dissimilarity matrix)
sil <- tryCatch(
    expr= { 
        cluster::silhouette(
            x =as.integer(clusters),   # integer vector of the initial cluster
            dmatrix=as.matrix(cluster::daisy(mtchs)))  # distance matrix
        },
    error=function(e) { return(NULL) }
)


# Save the silhouette object
saveRDS(sil, paste0(results_path, "/silhouette_", res, ".rds"))

# Finally, calculate grouped metrics
message(paste0("Grouping ", res, "..."))


# Compute group scores by passing a data table and clusters
grp <- tryCatch(
    expr={ group_scores(as.data.table(mtchs), clusters) },
    error=function(e) { NULL }
    )


# Ocassionally the sil is returned NA by the function cluster::silhouette() when
# - the size of input dataset is small and/or
# - the resolution is too high
# tryCatch() is applied in order to avoid possible errors under this condition
sil <- tryCatch(

    # Code to run (will return a data frame containing cluster, avg_sil, and res)
    expr = { group_sil(sil, res) },

    # If sil is returned NA by cluster::silhouette(),
    # an empty data frame is created
    error = function(e) {
        message("\n\n")
        message("Computing sil has failed due to missing sil!")
        message("This returns an empty data frame.")
        return(data.frame())
    }
    )


```

## Save objects


```{r save}

saveRDS(grp, paste0(results_path, "/frequency_grouped_", res, ".rds"))   # short (seconds)
saveRDS(sil, paste0(results_path, "/silhouette_grouped_", res, ".rds"))  # short (seconds)
saveRDS(sampled.obj, paste0(results_path, "/seurat_obj_", res, ".rds"))   # takes nearly 20min


```



## Session info

```{r session_info}

sessionInfo()

```


