---
title: "Cluster Seurat"
output:
    html_document:
            toc: true
            toc_float: true
            code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    warning=FALSE,
    message=FALSE
    )

```

```{r packages_and_functions, include=FALSE}
library(tidyverse)
library(Signac)
library(Seurat)
library(knitr)
library(future)
set.seed(1234)
```

```{r parallelize, include=FALSE}

# Assign the number of cores for parallelization
ncpus <- future::availableCores()

# Adjust max memory allowed
options(future.globals.maxSize=40000 * 1024^2,
        mc.cores=ncpus)

# Parallelize the run
plan("multicore", workers=ncpus)
```

# Set variables

We'll get all the sample and replicate specific variables listed in samples.tsv and units.tsv, respectively. 
We'll also set paths to input files.

```{r variables}
obj_input <- params[['seurat_in']]
choice_input <- params[['res_choice']]
snakemake.config <- yaml::read_yaml(params[['config']])
comm_detection <- snakemake.config[['cluster']][['detection_method']]
```


# Read in data

For input file paths listed in samples.tsv, we'll read in data from files in one of the following formats: 
'HDF5', 'MEX', or 'RDS'.

If multiple assays are present in the 'HDF5' or 'RDS' files, a list of sparse matrices ('dgCMatrix') 
will be created. 'read_10X_mm()' mimics Seurat::Read_10X() but allows import of multiple MEX matrices.

```{r input}

# Import Seurat obj
obj <- readRDS(obj_input)

# Set resolution to the previously computed (by chooser) or defined value
res_use <- readRDS(choice_input)

# Convert to numeric of the resolution is still a character obj
if (!is.numeric(res_use)) {
    res_use <- as.numeric(res_use)

}

print('Input Seurat object:')
print(obj)

print(paste('Resolution to use for community detection:', res_use))

# Retrieve community detection algorithm
comm_detection <- ifelse(is.null(comm_detection), 2, comm_detection)
comm_label <- ifelse(comm_detection == 1, "Louvain", ifelse(
    comm_detection == 2, "Louvain with multilevel refinement", ifelse(
    comm_detection == 3, "Smart local moving (SLM)", ifelse(
    comm_detection == 4, "Leiden", "Undetermined"
))))

print(paste('Algorithm to use for community detection:', comm_label))
reference <- "Ludo Waltmana and Nees Jan van Eck, Eur. Phys. J. B (2013)"
paper_link <- "https://link.springer.com/article/10.1140/epjb/e2013-40829-0"
```

Refer to the following reference for community detection algorithms: [`r reference`](`r paper_link`)

```{r vars_cluster}

# Retrieve the configuration for the number of dimensions
npcs <- snakemake.config[['chooser']][['groups']][[ params[['seurat']] ]]$npcs

# Break if npcs returns NULL
if (is.null(npcs)) {
    msg <- paste0(
        "npcs missing in chooser[['groups']][['",
        params[['seurat']],
        "']]")
    stop(msg)
}

# Identify dimensionality reduction performed in the previous step for the input obj
dim_reductions <- c("pca", "lsi")
reduction <- names(obj@reductions)
reduction_idx <- sapply(dim_reductions, function(r) str_detect(reduction, r))

# Break if the number of dimensionality reduction is 0 or more than 1 unintegrated dimensionality reductions
# found
if (sum(reduction_idx) > 1) {
    if (any(str_detect(reduction, "integrated"))) {
        reduction <- reduction[str_detect(reduction, "integrated")]
    }
    else {
        stop("More than one dimensionality reduction found!")
    }
} else if (sum(reduction_idx) == 0) {
    stop("Dimensionality reduction missing!")
}

# Identify assay in the current obj
assay <- obj[[reduction]]@assay.used

# Find the current assay in use
assay <- obj[[reduction]]@assay.used

# Determine the number of minimum dimensions
if (grepl("lsi", reduction)) {
    npcs_min <- 2
} else {
    npcs_min <- 1
}

# uwot fails with less that 50 cells
# https://github.com/satijalab/seurat/issues/4312
# if ( nrow(obj@reductions[[reduction]]) >= 50 ) {
#     umap_method <- "uwot"
# } else {
#     umap_method <- "umap-learn"
# }

umap_method <- "uwot"
```

# Clustering

Cells are clustered using the `FindNeighbors` followed by `FindClusters` functions in Seurat.

```{r cluster}

# Assign a name of metadata column for new cluster (e.g. "integrated_1_snn_res.0.8")
cluster_colname <- paste0(params[['seurat']],
                          paste0('_snn_res.', res_use))

print("Calling FindNeighbors...")
obj <- FindNeighbors(
    object = obj,
    reduction = reduction,
    dims = npcs_min:npcs
)

print("Calling FindClusters...")
obj <- FindClusters(
    object = obj,
    algorithm = comm_detection,
    resolution = res_use
)

print('Finished finding neighbors for this assay.')

# Ensure that the cluster is given correct name in the metadata
obj@meta.data[[cluster_colname]] <- obj$seurat_clusters


print(paste(cluster_colname, "column added to metadata for the current clustering"))
```

# Visiualize {.tabset}


```{r umap}

# Run UMAP
obj <- tryCatch(
    expr={
        RunUMAP(obj,
                dims = npcs_min:npcs,
                reduction = reduction,
                assay = assay,
                umap.method = umap_method)
    },
    error=function(e) {
        umap_method <- "umap-learn"
        RunUMAP(obj,
                dims = npcs_min:npcs,
                reduction = reduction,
                assay = assay,
                umap.method = umap_method)
    },
    finally={
        print(paste(umap_method, "used for UMAP"))
    }
)
```

```{r visualization, results='asis'}

if ("umap" %in% names(obj@reductions)) {

    cat('\n\n## By orig.ident\n\n')
    plot <- DimPlot(
      obj,
      reduction = "umap",
      pt.size = 0.5,
      group.by = "orig.ident"
    )

    print(plot)

    cat('\n\n## By', cluster_colname, '\n\n')
    plot <- DimPlot(
      obj,
      reduction = "umap",
      pt.size = 0.5,
      group.by = cluster_colname
    )

    print(plot)
    cat('\n\n')
} else {
    cat("No UMAPs generated!")
}
```


# Save Seurat object

```{r save}
cat('Updated Seurat object:')
print(obj)

# save main seurat object
saveRDS(object=obj, file=params[['seurat_out']])
print(paste('Updated Seurat object saved in', params[['seurat_out']]))

# save barcodes for later use (combining multiple processed objects)
barcodes <- colnames(obj)
saveRDS(object=barcodes, file=params[['barcodes_out']])
print(paste('Barcodes saved in', params[['barcodes_out']]))
```


## Session Info

```{r session_info}
sessionInfo()

```



