---
title: "Cluster Seurat"
output:
    html_document:
            toc: true
            toc_float: true
            code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    warning=FALSE,
    message=FALSE
    )

```

```{r packages_and_functions, include=FALSE}
library(Signac)
library(Seurat)
library(knitr)
library(future)
set.seed(1234)
```

## Parallelization

```{r parallelize, include=FALSE}

# Assign the number of cores for parallelization
ncpus <- future::availableCores()

# Adjust max memory allowed
options(future.globals.maxSize=40000 * 1024^2,
        mc.cores=ncpus)

# Parallelize the run
plan("multicore", workers=ncpus)
```

```{r}
params = list(
      seurat_in = "results/normalize_reduce_dims/unintegrated_2/unintegrated_2.rds",
      config = "/vf/users/BSPC-Petros/multiome-wf/workflow/config/multiome-config/config.yaml",
      res_choice = "results/chooser_plot/unintegrated_2/choice.rds",
      seurat = "unintegrated_2",
      res_use = "0.6",
      seurat_out = "results/cluster/unintegrated_2/clustered_data.rds",
      barcodes_out = "results/cluster/unintegrated_2/barcodes.rds")
```


## Set variables

We'll get all the sample and replicate specific variables listed in samples.tsv and units.tsv, respectively. We'll also set paths to input files.

```{r variables}
obj_input <- params[['seurat_in']]
choice_input <- params[['res_choice']]
snakemake.config <- yaml::read_yaml(params[['config']])
```


## Read in data

For input file paths listed in samples.tsv, we'll read in data from files in one of the following formats: 'HDF5', 'MEX', or 'RDS'.
If multiple assays are present in the 'HDF5' or 'RDS' files, a list of sparse matrices ('dgCMatrix') will be created.
'read_10X_mm()' mimics Seurat::Read_10X() but allows import of multiple MEX matrices.

```{r input}

# Import Seurat obj
obj <- readRDS(obj_input)

# Set resolution to the previously computed (by chooser) or defined value
if (params[['res_use']] == "None") {
    res_use <- readRDS(choice_input)
} else {
    res_use <- params[['res_use']]
}

# Convert to numeric of the resolution is still a character obj
if (!is.numeric(res_use)) {
    res_use <- as.numeric(res_use)
}

print('Input Seurat object:')
print(obj)

print(paste('Resolution to use for community detection:', res_use))
```

```{r vars_cluster}

# Retrieve the configuration for the number of dimensions
npcs <- snakemake.config[['chooser']][['groups']][[ params[['seurat']] ]]$npcs

# Find the primary dimensionality reduction
targets <- c("pca", "lsi")
idx <- grep(paste(targets,collapse="|"), names(obj@reductions))
reduction <- names(obj@reductions)[idx]

# Find the current assay in use
assay <- obj[[reduction]]@assay.used

# Determine the number of minimum dimensions
if (grepl("lsi", reduction)) {
    npcs_min <- 2
} else {
    npcs_min <- 1
}

# uwot fails with less that 50 cells
# https://github.com/satijalab/seurat/issues/4312
# if ( nrow(obj@reductions[[reduction]]) >= 50 ) {
#     umap_method <- "uwot"
# } else {
#     umap_method <- "umap-learn"
# }

umap_method <- "uwot"
```

## Clustering

Cells are clustered using the `FindNeighbors` followed by `FindClusters` functions in Seurat.

```{r cluster}

# Assign a name of metadata column for new cluster (e.g. "integrated_1_snn_res.0.8")
cluster_colname <- paste0(params[['seurat']],
                          paste0('_snn_res.', res_use))

print("Calling FindNeighbors...")
obj <- FindNeighbors(
    object = obj,
    reduction = reduction,
    dims = npcs_min:npcs
)

# not really needed not that weighted nn is in new rule, delete later
# graph_use = paste0(assay_use, '_snn')
# print(graph_use)

print("Calling FindClusters...")
obj <- FindClusters(
  object = obj,
  resolution = res_use
)

print('Finished finding neighbors for this assay.')

# Ensure that the cluster is given correct name in the metadata
obj@meta.data[[cluster_colname]] <- obj$seurat_clusters


print(paste(cluster_colname, "column added to metadata for the current clustering"))
```

## UMAP

```{r umap}

# Run UMAP
obj <- tryCatch(
    expr={
        RunUMAP(obj,
                dims = npcs_min:npcs,
                reduction = reduction,
                assay = assay,
                umap.method = umap_method)
    },
    error=function(e) {
        umap_method <- "umap-learn"
        RunUMAP(obj,
                dims = npcs_min:npcs,
                reduction = reduction,
                assay = assay,
                umap.method = umap_method)
    },
    finally={
        print(paste(umap_method, "used for UMAP"))
    }
)
```


## Visiualize {.tabset}

```{r visualization, results='asis'}

cat('### By orig.ident')
plot <- DimPlot(
  obj,
  reduction = "umap",
  pt.size = 0.5,
  group.by = "orig.ident"
)

print(plot)
cat('\n\n')

cat('### By', cluster_colname)
plot <- DimPlot(
  obj,
  reduction = "umap",
  pt.size = 0.5,
  group.by = cluster_colname
)

print(plot)
```


## Save Seurat object

```{r save, eval=FALSE}

cat('Updated Seurat object:')
print(obj)

# save main seurat object
saveRDS(object=obj, file=params[['seurat_out']])
print(paste('Updated Seurat object saved in', params[['seurat_out']]))

# save barcodes for later use (combining multiple processed objects)
barcodes <- colnames(obj)
saveRDS(object=barcodes, file=params[['barcodes_out']])
print(paste('Barcodes saved in', params[['barcodes_out']]))
```


## Session Info

```{r session_info}
sessionInfo()

```



