---
title: "Cluster Seurat"
output:
    html_document:
            keep_md: yes
            toc: true
            toc_float: true
            code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    warning=FALSE,
    message=FALSE,
    results='hide'
    )

```

```{r packages_and_functions, include=FALSE}
library(Signac)
library(Seurat)
library(knitr)
library(future)
set.seed(1234)
```

## Parallelization

```{r parallelize, include=FALSE}

# Assign the number of cores for parallelization
ncpus <- future::availableCores()

# Adjust max memory allowed
options(future.globals.maxSize=40000 * 1024^2,
        mc.cores=ncpus)

# Parallelize the run
plan("multicore", workers=ncpus)


```





## Set variables

We'll get all the sample and replicate specific variables listed in samples.tsv and units.tsv, respectively. We'll also set paths to input files.

```{r variables}
obj_input = params[['seurat_in']]
choice_input = params[['res_choice']]
snakemake.config <- yaml::read_yaml(params[['config']])
```


## Read in data

For input file paths listed in samples.tsv, we'll read in data from files in one of the following formats: 'HDF5', 'MEX', or 'RDS'.
If multiple assays are present in the 'HDF5' or 'RDS' files, a list of sparse matrices ('dgCMatrix') will be created.
'read_10X_mm()' mimics Seurat::Read_10X() but allows import of multiple MEX matrices.

```{r input}
obj = readRDS(obj_input)
res_use = readRDS(choice_input)

cat('Input Seurat object:')
print(obj)

cat('Resolution to use for community detection:')
print(res_use)
```


## 

```{r vars_cluster}
npcs <- snakemake.config[['chooser']][['groups']][[ params[['seurat']] ]]$npcs

targets <- c("pca", "lsi")
idx <- grep(paste(targets,collapse="|"), names(obj@reductions))
reduction <- names(obj@reductions)[idx]
assay <- obj[[reduction]]@assay.used

if (grepl("lsi", reduction)) {
  npcs_min = 2
} else {
  npcs_min = 1
}

# uwot fails with less that 50 cells
# https://github.com/satijalab/seurat/issues/4312
if (nrow(obj@reductions[[reduction]]) >= 50 ) {
  umap_method = "uwot"
} else {
  umap_method = "umap-learn"
}

```

## 

```{r cluster}

# Assign a name of metadata column for new cluster (e.g. "integrated_1_snn_res.0.8")
cluster_colname <- paste0(params[['seurat']],
                          paste0('_snn_res.', res_use))


print("Calling FindNeighbors.")
obj <- FindNeighbors(
  object = obj,
  reduction = reduction,
  dims = npcs_min:npcs
)

# not really needed not that weighted nn is in new rule, delete later
# graph_use = paste0(assay_use, '_snn')
# print(graph_use)

print("Calling FindClusters")
obj <- FindClusters(
  object = obj,
  resolution = res_use
)

print('Finished finding neighbors for this assay.')

# Ensure that the cluster is given correct name in the metadata
obj@meta.data[[cluster_colname]] <- obj$seurat_clusters

```

## Visiualize

```{r umap}
obj <- RunUMAP(obj,
               dims = npcs_min:npcs,
               reduction = reduction,
               assay = assay,
               umap.method = umap_method
)

plot <- DimPlot(
  obj,
  reduction = "umap",
  pt.size = 0.5,
  group.by = "orig.ident"
)

plot

plot <- DimPlot(
  obj,
  reduction = "umap",
  pt.size = 0.5,
  group.by = "seurat_clusters"
)

plot
```


## Save Seurat object

```{r save, results='asis'}
cat('Updated Seurat object:')
print(obj)

# save main seurat object
saveRDS(object=obj, file=params[['seurat_out']])
cat('Done saving seurat object.')

# save barcodes for later use (combining multiple processed objects)
barcodes = colnames(obj)
saveRDS(object=barcodes, file=params[['barcodes_out']])
cat('Done saving cell barcodes.')
```


## Session Info

```{r session_info}
sessionInfo()

```



