---
title: "Combine Processed Seurats"
output:
    html_document:
            keep_md: yes
            toc: true
            toc_float: true
            code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    warning=FALSE,
    message=FALSE,
    results='hide'
    )

```

```{r packages_and_functions, include=FALSE}
library(Signac)
library(Seurat)
library(knitr)
library(future)
set.seed(1234)

source('common.R')
```

## Parallelization

```{r parallelize, include=FALSE}

# Assign the number of cores for parallelization
ncpus <- future::availableCores()

# Adjust max memory allowed
options(future.globals.maxSize=40000 * 1024^2,
        mc.cores=ncpus)

# Parallelize the run
plan("multicore", workers=ncpus)
```






## Set variables

We'll get all the sample and replicate specific variables listed in samples.tsv and units.tsv, respectively. We'll also set paths to input files.

```{r variables}
seurat_paths = unlist(strsplit(params[['seurats']], " "))
    
config_path = params[["config"]]

seurat_out = params[["seurat_out"]]
```


## Read in data

For input file paths listed in samples.tsv, we'll read in data from files in one of the following formats: 'HDF5', 'MEX', or 'RDS'.
If multiple assays are present in the 'HDF5' or 'RDS' files, a list of sparse matrices ('dgCMatrix') will be created.
'read_10X_mm()' mimics Seurat::Read_10X() but allows import of multiple MEX matrices.

```{r config_input}
config = yaml::read_yaml(config_path)

groups <- names(config$normalize$groups)




assay_list = sapply(seq_along(groups), function(x){
  group = groups[x]
  assay_c = config$normalize$groups[[group]]$assay_name
  return(assay_c)
})

norm_list = sapply(seq_along(groups), function(x){
  group = groups[x]
  norm_c = config$normalize$groups[[group]]$norm_method
  return(norm_c)
})

if (config$integrate$activate) {
  groups_int = names(config$integrate$groups)

  assay_list_int = rep("integrate", length(groups_int))
  
  norm_list_int = sapply(seq_along(groups_int), function(x){
    group = groups_int[x]
    norm_c = config$integrate$groups[[group]]$norm_method
    return(norm_c)
  })
  
  groups = c(groups, groups_int)
  assay_list = c(assay_list, assay_list_int)
  norm_list = c(norm_list, norm_list_int)
  
  rm(groups_int, assay_list_int, norm_list_int)
}

if (config$coembed$activate) {
  # fix later
}

if (config$weighted_nn$activate) {
  groups_wnn = names(config$weighted_nn$groups)

  assay_list_wnn = rep("wnn", length(groups_wnn))
  
  norm_list_wnn = rep("wnn", length(groups_wnn))
  
  groups = c(groups, groups_wnn)
  assay_list = c(assay_list, assay_list_wnn)
  norm_list = c(norm_list, norm_list_wnn)
  
  rm(groups_wnn, assay_list_wnn, norm_list_wnn)
}

# Select correct analysis depending on macs2 setting
if (config$macs == 'N') {
    index <- assay_list != "MACS"
    groups <- groups[index]
    assay_list <- assay_list[index]
    norm_list <- norm_list[index]
}





# Raise an error if macs2 configuration is incorrect
if (!(config$macs %in% c("Y", "N"))) {
    stop("macs2 is incorrectly configured. Double-check macs in the config.yaml file.")
}



groups
assay_list
norm_list
```

```{r input}

# fix later:
# export barcode list from all seurat objects output 
# from rules "cluster" and "weighted_nn"
# but for testing, rading in each object to get barcodes is OK
barcodes_list = lapply(seq_along(seurat_paths), function(x) {
  # current_path = seurat_paths[x]
  # obj = readRDS(current_path)
  # barcodes = colnames(obj)
  
  current_path = seurat_paths[x]
  barcodes = readRDS(file.path(dirname(current_path), "barcodes.rds"))
  return(barcodes)
})

barcodes_list = lapply(barcodes_list, sort)

# check which objs have the same barcodes
bc_counts = unique(
  sapply(barcodes_list, length)
)

seurat_list = list()
list_n = 1

for (x in seq_along(bc_counts)) {
  bc_count = bc_counts[x]
  
  cat("\ncombining objects with the following barcode counts")
  print(bc_count)
  # get an index of which seurat objects has the same # barcodes
  idx = which(sapply(barcodes_list, length) == bc_count)
  # subset into current barcode #s
  bc_sub = barcodes_list[idx]
  
  paths_sub = seurat_paths[idx]
  groups_sub = groups[idx]
  assay_sub = assay_list[idx]
  norm_sub = norm_list[idx]
  
  cat(paths_sub, groups_sub, assay_sub, norm_sub)
  
  seq_sub = seq(1:length(paths_sub))
  
  # In bc vecotrs of equal length, are call constituent bc's identiecal too?
  if ( all(sapply(X = bc_sub[-1], FUN = identical, bc_sub[[1]])) ) {
    
    idys = grep("sct", norm_sub)
    if (length(idys) > 0) {
      idy = idys[1]
    } else {
      idy = 1
    }
    
    # if an object contains sct, use that first
    paths_c = paths_sub[idy]
    groups_c = groups_sub[idy]
    assay_c = assay_sub[idy]
    norm_c = norm_sub[idy]
    
    cat(paths_c, groups_c, assay_c, norm_c)
    
    seurat_main = readRDS(paths_c)
    
    #update dim reduction names before mergino objects
    for (r in seq_along(seurat_main@reductions)) {
      dr_name = paste(
        seurat_main@reductions[[r]]@assay.used,
        names(seurat_main@reductions)[r],
        sep = "_")
      names(seurat_main@reductions)[r] = dr_name
      Key(object = seurat_main[[dr_name]]) = dr_name
    }
    
    colnames_use = colnames(seurat_main)
    
    # if any remaining assays exist, add to seurat object
    if (length(seq_sub) > 1) {
      # idy: either first object containing sct assay (if present) or first object
      remaining = seq_sub[-idy]
      
      paths_r = paths_sub[remaining]
      groups_r = groups_sub[remaining]
      assay_r = assay_sub[remaining]
      norm_r = norm_sub[remaining]
      
      cat(paths_r, groups_r, assay_r, norm_r)
      
      for (i in seq_along(remaining)) {
        # i=2 #testing
        paths_c = paths_r[i]
        groups_c = groups_r[i]
        assay_c = assay_r[i]
        norm_c = norm_r[i]
        
        cat(paths_c, groups_c, assay_c, norm_c)
        
        
        seurat_add = readRDS(paths_c)
        
        if (any(grepl("integrated", names(seurat_add@reductions)))) {
          # integrating embedding of peaks does not rename assay to "integrate"
          if (seurat_add@active.assay == "integrated") {
            assay_new = groups_c
          } else {
            # fix later:
            # possibly use 'make.names(groups_c)'
            assay_new = paste(groups_c,
                              seurat_add@active.assay,
                              sep = "_")
          }
        } else {
          assay_new = seurat_add@active.assay
          # assay_new = paste(seurat_add@active.assay,
          #                   groups_c,
          #                   sep = "_")
        }
        
        # do not transfer assay if object is "wnn", which only adds Dim Reducs
        if (!any(names(seurat_add@reductions) %in% "wnn_umap")) {
          if (any(names(seurat_main@assays) %in% assay_new)) {
            # if assays exist, overwrite (to keep nomalized and scaled slots)
            seurat_main[[seurat_add@active.assay]] = NULL
            seurat_main[[assay_new]] = seurat_add[[seurat_add@active.assay]]
          } else {
            # if not exist, just add new assay
            seurat_main[[assay_new]] = seurat_add[[seurat_add@active.assay]]
          }
        }
        
        if (any(grepl("integrated", names(seurat_add@reductions)))) {
          Key(seurat_main[[assay_new]]) = assay_new
        }
        
        # if object contains wnn, remove all other dim reductions,
        # they are dups from other objects
        if ( any(grepl("wnn_umap", names(seurat_add@reductions))) ) {
          for (j in rev(seq_along(names(seurat_add@reductions)))) {
            dr_c = names(seurat_add@reductions)[j]
            print(dr_c)
            if (!dr_c == "wnn_umap") {
              seurat_add@reductions[[dr_c]] = NULL
            }
          }
        }
        
        for (r in seq_along(seurat_add@reductions)) {
          if (any(grepl("integrated", names(seurat_add@reductions))) && 
              names(seurat_add@reductions)[r] != "umap") {
            dr_name = gsub("integrated", groups_c, names(seurat_add@reductions)[r])
            names(seurat_add@reductions)[r] = dr_name
            Key(object = seurat_add[[dr_name]]) = dr_name
            # next line is needed for transferring to new seurat assay
            seurat_add@reductions[[r]]@assay.used = assay_new
            
          } else if (any(grepl("integrated", names(seurat_add@reductions))) && 
                     names(seurat_add@reductions)[r] == "umap") {
            dr_name = paste(
              groups_c,
              names(seurat_add@reductions)[r],
              sep = "_")
            names(seurat_add@reductions)[r] = dr_name
            Key(object = seurat_add[[dr_name]]) = dr_name
            seurat_add@reductions[[r]]@assay.used = assay_new
            
          } else if (names(seurat_add@reductions)[r] == "wnn_umap") {
            dr_name = gsub("wnn", groups_c, names(seurat_add@reductions)[r])
            names(seurat_add@reductions)[r] = dr_name
            Key(object = seurat_add[[dr_name]]) = dr_name
          } else {
            dr_name = paste(
              seurat_add@reductions[[r]]@assay.used,
              names(seurat_add@reductions)[r],
              sep = "_")
            names(seurat_add@reductions)[r] = dr_name
            Key(object = seurat_add[[dr_name]]) = dr_name
          }
          
          # dr_transfer = names(seurat_add@reductions)[r]
          seurat_main[[dr_name]] = seurat_add[[dr_name]]
        }
        
        # transfer clustering metadata
        ident_pattern = "snn_res\\.[0-9]+.*"
        idc = grep(ident_pattern, names(seurat_add@meta.data))
        if (length(idc) > 0) {
            clust_column = names(seurat_add@meta.data)[idc]
            if (any(grepl("wsnn", clust_column))) {
                idw = which(grepl("wsnn", clust_column))
                clust_column = clust_column[idw]
                col_name_new = gsub("wsnn", groups_c, clust_column)
            } else if (grepl("integrated", clust_column)) {
                col_name_new = gsub("integrated", groups_c, clust_column)
            } else {
              col_name_new = clust_column
            }
        }
        
		    # Re-order seurat_add metadata rows based on seurat_main colnames
        rows.add <- colnames_use
        meta.add <- seurat_add@meta.data[rows.add, ]
        seurat_main = AddMetaData(
            object = seurat_main, 
            metadata = meta.add[,clust_column], 
            col.name = col_name_new
        )

      }
    }
  } else {
    #fix later:
    # do something if 2 seurat objects have equal # barcodes, bc's not identical
    #sapply(X = bc_sub[-1], FUN = identical, bc_sub[1])
  }
  seurat_list[[list_n]] = seurat_main
  list_n = list_n + 1
}

```



## Save Seurat object

```{r save, results='asis'}
cat('Updated Seurat object:')
print(seurat_list)

saveRDS(object=seurat_list, file=seurat_out)
cat('Done saving seurat object.')
```


## Session Info

```{r session_info}
sessionInfo()

```


