---
title: "Combine Processed Seurats"
output:
    html_document:
            toc: true
            toc_float: true
            code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    warning=FALSE,
    message=FALSE
    )

```

```{r packages_and_functions, include=FALSE}
library(tidyverse)
library(parallel)
library(Signac)
library(Seurat)
library(knitr)
library(future)
set.seed(1234)

source('common.R')
```


```{r parallelize, include=FALSE}

# Assign the number of cores for parallelization
ncpus <- future::availableCores()

# Adjust max memory allowed
options(future.globals.maxSize=40000 * 1024^2,
        mc.cores=ncpus)

# Parallelize the run
plan("multicore", workers=ncpus)
```

```{r}
params = list(
                seurats = "results/cluster/unintegrated_0/clustered_data.rds results/cluster/unintegrated_1/clustered_data.rds results/cluster/unintegrated_2/clustered_data.rds results/cluster/unintegrated_3/clustered_data.rds results/cluster/integrated_0/clustered_data.rds results/cluster/integrated_1/clustered_data.rds results/cluster/integrated_3/clustered_data.rds results/wnn/wnn_0/clustered_data.rds results/wnn/wnn_1/clustered_data.rds",
                config = "/vf/users/BSPC-Petros/multiome-wf/workflow/config/multiome-config/config.yaml",
                seurat_out = "results/combine/seurat_combined.rds" )
```


# Set variables

We'll get all the sample and replicate specific variables listed in samples.tsv and units.tsv, 
respectively. We'll also set paths to input files.

```{r variables}
# Retrieve input/output file paths
seurat_paths <- unlist(strsplit(params[['seurats']], " "))
config_path <- params[["config"]]
seurat_out <- params[["seurat_out"]]
```


# Read in data

For input file paths listed in samples.tsv, we'll read in data from files in one of the following formats: 
'HDF5', 'MEX', or 'RDS'.

If multiple assays are present in the 'HDF5' or 'RDS' files, a list of sparse matrices ('dgCMatrix') 
will be created. 'read_10X_mm()' mimics Seurat::Read_10X() but allows import of multiple MEX matrices.

```{r config_input}

# Import user configuration
config <- yaml::read_yaml(config_path)

# Retrieve user-configured groups
groups <- names(config$normalize$groups)
assay_list <- sapply(groups, function(group) { 
    config$normalize$groups[[group]]$assay_name
})

# Retrieve user-configured normalizations across the groups
norm_list <- sapply(groups, function(x){
    config$normalize$groups[[x]]$norm_method
})

# If integration is enabled
if (config$integrate$activate) {
    # Create a vector storing the names for integrated groups
    groups_int <- names(config$integrate$groups)

    # Add a placeholder for integrated groups
    assay_list_int <- rep("integrate", length(groups_int))

    # Create a vector storing normalization method for each integrated group
    norm_list_int <- sapply(groups_int, function(x){
        config$integrate$groups[[x]]$norm_method
    })

    # Create a vector storing all group names
    groups <- unique(c(groups, groups_int))

    # Add activated assays and normalizations 
    assay_list <- c(assay_list, assay_list_int)
    norm_list <- c(norm_list, norm_list_int)

    # Delete unnecessary variables
    rm(groups_int, assay_list_int, norm_list_int)
}

# If multimodal analysis is enabled
if (config$weighted_nn$activate) {
    # Retrieve multimodal group names
    groups_wnn <- names(config$weighted_nn$groups)

    # Add a placeholder for mulimodal groups
    assay_list_wnn <- rep("wnn", length(groups_wnn))
    norm_list_wnn <- rep("wnn", length(groups_wnn))

    # Add group, assay, and normalization names corresponding to multimodal analysis
    groups <- c(groups, groups_wnn)
    assay_list <- c(assay_list, assay_list_wnn)
    norm_list <- c(norm_list, norm_list_wnn)

    rm(groups_wnn, assay_list_wnn, norm_list_wnn)
}

# Select correct analysis depending on macs2 setting
if (config$macs$run == 'N') {
    index <- assay_list != "MACS"
    groups <- groups[index]
    assay_list <- assay_list[index]
    norm_list <- norm_list[index]
}


# Raise an error if macs2 configuration is incorrect
if (!(config$macs$run %in% c("Y", "N"))) {
    stop("macs2 is incorrectly configured. Double-check macs in the config.yaml file.")
}

print("Groups analyzed:")
groups

print("Assays analyzed:")
assay_list

print("Preprocessing performed:")
norm_list
```

```{r input_barcodes}

# Import saved barcodes
barcodes_list <- lapply(seurat_paths, function(x) {
    readRDS(file.path(dirname(x), "barcodes.rds")) %>%
        sort()
}) %>%
set_names(str_split_i(seurat_paths, "/", 3))

# check which objs have the same barcodes
bc_counts <- sapply(barcodes_list, length) %>%
    unique()

# Create a list of lists storing barcodes, paths to Seurat obj, groups,
# assays, and normalizations for corresponding barcode counts
bc_summary_list <- mclapply(bc_counts, function(bc_count) {

    # Get an index of which seurat objects has the same # barcodes
    idx <-sapply(barcodes_list, length) == bc_count

    # Return a list storing everything
    list(
        bc_sub=barcodes_list[idx],
        paths_sub=seurat_paths[idx],
        groups_sub=groups[idx],
        assay_sub=assay_list[idx],
        norm_sub=norm_list[idx]
    )
}) %>%
set_names(paste0("nBarcodes_", bc_counts))

print(paste(
    "Barcode counts per obj:",
    paste(names(bc_summary_list), collapse=", "))
)
```

```{r build_obj}

seurat_list <- lapply(names(bc_summary_list), function(name) {

    bc_count <- bc_summary_list[[name]]
    print(paste(name, "processed..."))
    print("---------------------------------------------------------------------------")

    # subset into current barcode #s
    bc_sub <- bc_count$bc_sub
    paths_sub <- bc_count$paths_sub
    groups_sub <- bc_count$groups_sub
    assay_sub <- bc_count$assay_sub
    norm_sub <- bc_count$norm_sub

    # In bc vecotrs of equal length, are they all identical barcodes?
    bc_identical <- sapply(
        bc_sub[-1], 
        function(bc) identical(bc, bc_sub[[1]])
        )

    # If barcodes are all identical,
    if ( all(bc_identical) ) {

        # Get indices for all SCT-normalized groups
        idys <- grep("sct", norm_sub)

        # Pick one index for SCT-normalized groups
        idy <- ifelse(length(idys) > 0, idys[1], 1)

        # if an object contains sct, use that first
        paths_c <- paths_sub[idy]
        groups_c <- groups_sub[idy]
        assay_c <- assay_sub[idy]
        norm_c <- norm_sub[idy]

        print(paste("Path to Seurat obj:", paths_c))
        print(paste("Corresponding group:", groups_c))
        print(paste("Corresponding assay:", assay_c))
        print(paste("Corresponding preprocessing:", norm_c))

        print("---------------------------------------------------------------------------")

        # Import the main Seurat obj
        seurat_main <- readRDS(paths_c)

        # Update dim reduction names before merging objects
        for (r in seq_along(seurat_main@reductions)) {

            # Determine a new name
            # e.g. pca to SCT_pca, umap to SCT_umap
            dr_name <- paste(
                seurat_main@reductions[[r]]@assay.used,
                names(seurat_main@reductions)[r], sep = "_")

            names(seurat_main@reductions)[r] <- dr_name
            Key(object=seurat_main[[dr_name]]) <- dr_name
        }

        # If any remaining assays exist, add to seurat object
        if (length(paths_sub) > 1) {

            # idy: either first object containing sct assay (if present) or first object
            remaining <- seq_along(paths_sub)[-idy]

            paths_r <- paths_sub[remaining]
            groups_r <- groups_sub[remaining]
            assay_r <- assay_sub[remaining]
            norm_r <- norm_sub[remaining]

            # Loop over to add individual assays to the main Seurat object
            for (i in seq_along(remaining)) {

                paths_c <- paths_r[i]
                groups_c <- groups_r[i]
                assay_c <- assay_r[i]
                norm_c <- norm_r[i]

                print(paste("Path to Seurat obj:", paths_c))
                print(paste("Corresponding group:", groups_c))
                print(paste("Corresponding assay:", assay_c))
                print(paste("Corresponding preprocessing:", norm_c))

                print("---------------------------------------------------------------------------")

                seurat_add <- readRDS(paths_c)

                # If the current assay is integrated one
                if (any(grepl("integrated", names(seurat_add@reductions)))) {
                    # integrating embedding of peaks does not rename assay to "integrate"
                    assay_new <- ifelse(
                        seurat_add@active.assay == "integrated",
                        group_c,
                        paste(groups_c, seurat_add@active.assay, sep = "_")
                    )
                } else {
                    assay_new <- seurat_add@active.assay
                }

                # Do not transfer assay if object is "wnn", which only adds Dim Reducs
                if (!any(names(seurat_add@reductions) %in% "wnn_umap")) {
                    if (any(names(seurat_main@assays) %in% assay_new)) {
                        # if assays exist, overwrite (to keep nomalized and scaled slots)
                        seurat_main[[seurat_add@active.assay]] <- NULL
                        seurat_main[[assay_new]] <- seurat_add[[seurat_add@active.assay]]
                    } else {
                        # if not exist, just add new assay
                        seurat_main[[assay_new]] <- seurat_add[[seurat_add@active.assay]]
                    }
                }

                if (any(grepl("integrated", names(seurat_add@reductions)))) {
                    tryCatch( expr={ Key(seurat_main[[assay_new]]) <- assay_new },
                        error=function(e) {
                            # add a new assay if absent
                            seurat_main[[assay_new]] <- seurat_add[[seurat_add@active.assay]]
                        }
                    )
                }

                # if object contains wnn, remove all other dim reductions,
                # they are dups from other objects
                if (any(grepl("wnn_umap", names(seurat_add@reductions)))) {
                    for (j in rev(seq_along(names(seurat_add@reductions)))) {
                        dr_c <- names(seurat_add@reductions)[j]
                        # Remove non-wnn reductions if present
                        if (!dr_c == "wnn_umap") {
                            seurat_add@reductions[[dr_c]] <- NULL
                        }
                    }
                }

                for (r in seq_along(seurat_add@reductions)) {
                    reduction_name <- names(seurat_add@reductions)[r]
                    if (any(grepl("integrated", names(seurat_add@reductions))) && 
                        reduction_name != "umap") {
                        dr_name <- gsub("integrated", groups_c, reduction_name)
                        names(seurat_add@reductions)[r] <- dr_name
                        Key(object = seurat_add[[dr_name]]) <- dr_name
                        # next line is needed for transferring to new seurat assay
                        seurat_add@reductions[[r]]@assay.used <- assay_new

                    } else if (any(grepl("integrated", names(seurat_add@reductions))) && 
                        reduction_name == "umap") {
                            dr_name <- paste(groups_c,
                                             reduction_name,
                                             sep = "_")
                            names(seurat_add@reductions)[r] <- dr_name
                            Key(object = seurat_add[[dr_name]]) <- dr_name
                            seurat_add@reductions[[r]]@assay.used <- assay_new

                    } else if (reduction_name == "wnn_umap") {
                        dr_name <- gsub("wnn", groups_c, reduction_name)
                        names(seurat_add@reductions)[r] <- dr_name
                        Key(object = seurat_add[[dr_name]]) <- dr_name
                    } else {
                        dr_name <- paste(
                            seurat_add@reductions[[r]]@assay.used,
                            reduction_name,
                            sep = "_"
                            )
                        names(seurat_add@reductions)[r] <- dr_name
                        Key(object = seurat_add[[dr_name]]) <- dr_name
                    }

                    seurat_main[[dr_name]] <- seurat_add[[dr_name]]
                }

                # Transfer clustering metadata
                ident_pattern <- "snn_res\\.[0-9]+.*"
                idc <- grep(ident_pattern, names(seurat_add@meta.data))
                if (length(idc) > 0) {
                    clust_column <- names(seurat_add@meta.data)[idc]
                    col_name_new <- if (any(grepl("wsnn", clust_column))) {
                        idw <- which(grepl("wsnn", clust_column))
                        clust_column <- clust_column[idw]
                        sub("wsnn", groups_c, clust_column)
                    } else {
                        clust_column
                    }
                }

                # Re-order seurat_add metadata rows based on seurat_main colnames
                meta.add <- seurat_add@meta.data[colnames(seurat_main), ]
                seurat_main <- AddMetaData(
                    object = seurat_main, 
                    metadata = meta.add[,clust_column], 
                    col.name = col_name_new
                    )
            }
        }
        seurat_main

    } else {
        NULL
    #fix later:
    # do something if 2 seurat objects have equal # barcodes, bc's not identical
    #sapply(X = bc_sub[-1], FUN = identical, bc_sub[1])
    }
}) %>%
set_names(names(bc_summary_list))

# Remove NULL object from the list
seurat_list <- seurat_list[sapply(seurat_list, function(o) !is.null(o))]
```



# Save Seurat object

```{r save, results='asis'}

print('Updated Seurat obj:')
print(seurat_list)

# saveRDS(object=seurat_list, file=seurat_out)
print(paste("Seurat obj saved to", seurat_out))
```


## Session Info

```{r session_info}
sessionInfo()

```


