---
title: "Creating Seurat"
output:
    html_document:
            toc: true
            toc_float: true
            code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    warning=FALSE,
    message=FALSE
    )
```

```{r packages_and_functions, include=FALSE}
library(dplyr)
library(purrr)
library(Signac)
library(Seurat)
library(GenomeInfoDb)
library(knitr)
library(future)
set.seed(1234)

source('common.R')
```

## Parallelization

```{r parallelize, include=FALSE}

# Assign the number of cores for parallelization
ncpus <- future::availableCores()

# Adjust max memory allowed
options(future.globals.maxSize=40000 * 1024^2,
        mc.cores=ncpus)

# Parallelize the run
plan("multicore", workers=ncpus)


```

## Set variables

We'll get all the sample and replicate specific variables listed in samples.tsv and units.tsv, respectively. We'll also set paths to input files.

```{r variables}

# input matrix paths
tenx_features <- unlist(strsplit(params[['tenx_type']], " "))
tenx_inputs <- unlist(strsplit(params[['tenx']], " "))

print(paste("Loading the following multi-assay or single-assay matrices:", tenx_features))


# get sample/replicate info, and load all tsv files
wc_sample <- params[['wc_sample']]
wc_replicate <- params[['wc_replicate']]

# sample.tsv info
sample_info <- readr::read_delim(params[['sample_tsv']], 
                          delim = "\t", escape_double = FALSE, 
                          trim_ws = TRUE)
# select row that match sample and replicate info
if (wc_replicate == "nan") {
    idx <- sample_info[['sample']] == wc_sample
} else {
    idx <- sample_info[['sample']] == wc_sample & 
                              sample_info[['replicate']] == wc_replicate
}
sample_info <- sample_info[idx, ]
print("Sample table:")
print(as.data.frame(sample_info))

# aggregates.tsv info
if (params[['aggregates_tsv']] != "") {
    aggregates_info <- readr::read_delim(params[['aggregates_tsv']],
                                         delim = "\t", escape_double = FALSE,
                                         trim_ws = TRUE)
    
    cols <- c("sample", "replicate", "library_id", "bc_suffix", "metadata1")
    if (nrow(aggregates_info) == 0) {
        print("Empty aggregates.tsv file found!")
        print("Please fix or delete aggregates.tsv")
    } else if ( any(!colnames(aggregates_info) %in% cols) ) {
        # fix later, add better handling if tsv has incorrect header
        print("aggregates.tsv file has incorrect columns!")
        print("aggregates.tsv requires the following conumn names:")
        print(cols)
        print("Please fix aggregates.tsv")
    } else {
        # select row that has sample and replicate info
        if (wc_replicate == "nan") {
            idx <- aggregates_info[['sample']] == wc_sample
        } else {
            idx <- aggregates_info[['sample']] == wc_sample & 
                aggregates_info[['replicate']] == wc_replicate
        }
        aggregates_info <- aggregates_info[idx, ]
        print("Aggregates table:")
        print(as.data.frame(aggregates_info))
    }
}

# assays.tsv info
if (params[['assays_tsv']] != "") {
    assays_info <- readr::read_delim(params[['assays_tsv']],
                                     delim = "\t", escape_double = FALSE,
                                     trim_ws = TRUE)

    cols <- c("sample", "replicate", "assay.name", "path")
    if (nrow(assays_info) == 0) {
        print("Empty assays.tsv file found!")
        print("Please fix or delete assays.tsv")
        colnames(assays_info) = cols
    } else if ( any(!colnames(assays_info) %in% cols) ) {
        # fix later, add better handling if tsv has incorrect header
        print("assays.tsv file has incorrect columns!")
        print("assays.tsv requires the following conumn names:")
        print(cols)
        print("Please fix assays.tsv")
    } else {
        # select row that has sample and replicate info
        if (wc_replicate == "nan") {
            idx <- assays_info[['sample']] == wc_sample
        } else {
            idx <- assays_info[['sample']] == wc_sample & 
                assays_info[['replicate']] == wc_replicate
        }
        assays_info <- assays_info[idx, ]
        assays_info <- na.omit(assays_info)
    }

    # append custom count matrices if present
    if (nrow(assays_info) > 0) {
        tenx_features <- c(tenx_features, assays_info[['assay.name']])
        tenx_inputs <- c(tenx_inputs, assays_info[['path']])
    }
}


# load annotation file
annotation <- readRDS(params[['annotation']])
print('Annotation object:')
print(annotation)
```


## Read in data

For input file paths listed in samples.tsv, we'll read in data from files in one of the following formats: 'HDF5', 'MEX', or 'RDS'.
If multiple assays are present in the 'HDF5' or 'RDS' files, a list of sparse matrices ('dgCMatrix') will be created.
'read_10X_mm()' mimics Seurat::Read_10X() but allows import of multiple MEX matrices.

```{r input}
# fix later, replace reading in entire data set with reading in 
# matrix feature names using hdf5r
tenx_data <- lapply(seq_along(tenx_features), function(x) {
    tenx_feature <- tenx_features[x]
    tenx_input <- tenx_inputs[x]
    
    print(paste('Reading in', tenx_feature, 'matrix from path:', tenx_input))
    if (grepl('.h5', tenx_input)) {
        tenx_data <- Seurat::Read10X_h5(filename = tenx_input)
    } else if (grepl('.rds', tenx_input)) {
        tenx_data <- readRDS(tenx_input)
    } else {
        tenx_data <- read_10X_mm(
            data.dir = tenx_input, 
            feature_list = tenx_feature
        )
    }
    
    if (class(tenx_data) != 'list') {
        tenx_data <- list(tenx_data)
    }
    
    # assumes reading in only one 10X matrix that is coerced by Seurat 
    # into sparse matrix
    if (is.null(names(tenx_data)) || length(names(tenx_data)) == 0) {
        if (length(tenx_data) == 1) {
            if (tenx_feature == "HDF5_Multiple_Assays" || tenx_feature == "RDS_Multiple_Assays") {
                cat("\nOne of the files in either HDF5 or RDS columns of 'samples.tsv' only contains a single cell-by-feature matrix!\n\n")
                cat("Files listed in HDF5 and RDS columns of 'samples.tsv' are for multi-assay files!\n")
                cat("HDF5 file heirarchy must use 10X Genomics cellranger compatible format.\n")
                cat("RDS must be a named list of 2 or more 'dgCMatrix' objects, list names corresponding to assay type.\n\n")
                cat("If cellranger (or similar program) generated the single assay cell-by-feature matrix, please move to any of the following columns in 'samples.tsv': 'Gene.Express', 'Peaks' and/or 'TF' or add to 'assays.tsv'.\n\n")
                cat("Quiting.\n")
                q(save = "no")
            } else {
                names(tenx_data) <- tenx_feature
            }
        }
    }

    return(tenx_data)
})

# some list elements will be nested, since Read10X_h5 or read_10x_mm will 
# return list if > 1 assay found
tenx_data <- unlist(tenx_data, recursive = FALSE)

# assay names must be unique per sample
# e.g. can't read in RNA mat from HDF5 and Gene.Expression cols in 'samples.tsv'
if (any(duplicated(names(tenx_data)))) {
    print("Can't have input matrices for the same assay list in multiple columns of 'samples.tsv'!")
    print("Found duplicate input matrices for the following assays:\n")
    print(names(tenx_data)[duplicated(names(tenx_data))])
    print("Please update the input files listed in 'samples.tsv' and 'assays.tsv'.")
    stop(save = "no")
}

# print(str(tenx_data))

# # Convert 10X Genomics 'Gene Expression' feature type to 'RNA'
# # Many Seurat functions look specifically for 'RNA' string in assays
# names(tenx_data)[names(tenx_data) %in% 'Gene Expression'] = 'RNA'
names(tenx_data) <- make.names(names(tenx_data))

# remove any prefixes added to 10X matrix barcodes upon loading by 
# Seurat (this isexpected behavior)
tenx_data <-  lapply(seq_along(tenx_data), function(x) {
    m_in <- tenx_data[[x]]
    old_names <- colnames(m_in)
    new_names <- gsub('^.*_', '', old_names)
    colnames(m_in) <- new_names
    return(m_in)
}) %>%
set_names(names(tenx_data))

# cat('list names and cell names were updated!')
# print(str(tenx_data))
```


## Ceate a list of Seurat Assay Objects

We'll convert the list of sparse matrices that was just read into memory and convert each list element into a Seurat 'Assay' object or Signac 'ChromatinAssay' object. The motivation for this is support for an arbitrary number of assays specified in the input/samples.tsv files, promoting flexibility in multimodal analyses and/or supporting or new assays in the future.

```{r assays}
# create list of Seurat Assay objects
assay_list <- lapply(seq_along(tenx_data), function(x) {
    feature_name = names(tenx_data)[x]
    m_in = tenx_data[[x]]
    # If a table of tagmentation loci (with coordinates and de-duplicated counts) is available,
    if (feature_name == 'Peaks' && sample_info[['fragments']] != 'None') {
        # create an obj for ZINBA 'Peaks' matrix
        obj <- Signac::CreateChromatinAssay(
            counts = m_in,
            sep = c(":", "-"),
            fragments = sample_info[['fragments']],
            annotation = annotation,
            min.features=-0.1
            )

    # If a table of tagmentation loci (with coordinates and de-duplicated counts) is unavailable,
    } else if (feature_name == 'Peaks' && sample_info[['fragments']] == 'None') {
        stop("Fragment file (fragments.tsv.gz) missing! Check your sample table.")
    } else {
        # Create an RNA obj
        obj <- Seurat::CreateAssayObject(counts = m_in)
    }
    return(obj)
}) %>%
set_names(names(tenx_data))
```


## Filter assay cell names

TEMPORARY - maybe keep as permanent feature after testing.

For test data sets, subsets of ATAC data may have many cells in which all peak counts are zero. 
This is due to the sparsity of scATAC data. This problem will likely arise in other assay/modality 
types such as CRISPR barcodes.

If sample assays have different cells/column names, creation of Seurat objects in next section breaks. 
So we'll detect cell names that are present in all assays and then select these columns for each assay.

```{r filter_cells}
print('Dimensions of assays before filtering:')
df <- as.data.frame(lapply(assay_list, dim))
rownames(df) <- c('Features', 'Cells')
print(df)

# Create a list storing cell barcodes in each modality
cells_keep <- lapply(seq_along(assay_list), function(x) {
    assay_use <- assay_list[[x]]
    cells <- colnames(assay_use)
    return(cells)
}) %>%
set_names(names(assay_list))

# Print the number of unique barcodes for each modality
print('Unique barcodes in each modality:')
lapply(cells_keep, length) %>% as.data.frame()

# Create a vector storing intersecting barcodes across all modalities
cells_keep <- Reduce('intersect', cells_keep)

print(paste('Number of cells to use that are common across all sample assays:', length(cells_keep)))

# Filter cells based on the intersection
assay_list <- lapply(seq_along(assay_list), function(x) {
    assay_use <- assay_list[[x]]
    assay_subset <- subset(assay_use, cells = cells_keep)
    return(assay_subset)
}) %>%
set_names(names(assay_list))

print('Dimensions of assays after filtering:')
df <- as.data.frame(lapply(assay_list, dim))
rownames(df) <- c('Features', 'Cells')
print(df)
```


## Ceate a Seurat object

We'll generate a Seurat object with the first element of the 'Assays' list, and then iterate over remaining assays, adding each to a new slot in Seurat object.

```{r seurat}
# create Seurat object with first assay from list
seurat <- Seurat::CreateSeuratObject(
       counts = assay_list[[1]],
       project = sample_info[['sample']],
       assay = names(assay_list)[1]
)
print("Initial Seurat obj created:")
print(seurat)

# if any remaining assays exist, add to seurat object
if (length(assay_list) > 1) {
    for (i in 2:length(assay_list)) {
        feature_name <- names(assay_list[i])
        print(paste("Assay added:", feature_name))
        a_in <- assay_list[[i]]
        seurat[[feature_name]] <- a_in
        print(seurat)
    }
}


# Add the sample info to barcodes
if (wc_replicate != "nan") {
    seurat <- RenameCells(
        object = seurat,
        add.cell.id = paste0(sample_info[['sample']], ".", sample_info[['replicate']])
    )
} else {
    seurat <- RenameCells(
        object = seurat,
        add.cell.id = sample_info[['sample']]
    )
}

# compute gene activities
# NOTE: 
# - this step aims to quantify the activity of each gene in the genome by assessing 
#   the chromatin accessibility associated with the gene
# - sums the fragments intersecting the gene body and promoter region
if (any(names(seurat@assays) %in% 'Peaks')) {
    DefaultAssay(seurat) <- 'Peaks'

    gene_activities <- Signac::GeneActivity(
        object = seurat,
        assay = 'Peaks'
    )
    print("Assay added: Gene.Activity")
    seurat[['Gene.Activity']] <- CreateAssayObject(counts = gene_activities)

    print('- adding Nucleosome Signal metadata')
    seurat <- NucleosomeSignal(seurat)
    print('- adding TSSEnrichment metadata')
    seurat <- TSSEnrichment(seurat)
    print(seurat)
}

# Stash mito QC stats, not informative unless whole cell GEX
if (any(names(seurat@assays) %in% 'Gene.Expression')) {
    mito_assay <- 'Gene.Expression'
} else if (any(names(seurat@assays) %in% 'Gene.Activity')) {
    mito_assay <- 'Gene.Activity'
}

# Add mito % to meta.data
seurat[["percent.mt"]] <- PercentageFeatureSet(
    object = seurat,
    pattern = "^[Mm][Tt]-",
    assay = mito_assay
)

print("Metadata columns present:")
print(colnames(seurat@meta.data))

# May need to add assay keys when creating Seurat objects from Assay object
# creating new seurat object or coercing from other formats (singlecellexperiment, etc)
#    assay names converted to all lower alphanumeric
# combining multiple seurat objects via merge() or integratedata()
#    assay names are kept as is after checking for pattern: ^[[:alnum:]]+_$ in assay key
# for now, keep behavior below, if needed later use assay names w/o converting to lower case.
key.df <- lapply(names(seurat@assays), function(assay_select) {
    print('Assay:')
    print(assay_select)
    print('Assay key:')
    key <- Key(object = seurat[[assay_select]])

    # Seurat names usually follow make.names() scheme, except for keys in assays
    # (i.e. seurat@assays$assay$key) which strips whitespace and '.'
    # Multiplexing Captue > Assay: Multiplexing.Capture > key: multiplexingcapture_

    if (length(key) == 0) {
        key <- paste0(tolower(assay_select), "_")
        Key(object = seurat[[assay_select]]) <- key
        print('Assay new key:')
        new_key <- Key(object = seurat[[assay_select]])
    }

    data.frame(Assay=assay_select, Key=key)
}) %>%
bind_rows()

print("Assay-Key pairs:")
print(key.df)
```

## Add Metadata to Seurat object

Lets add the following metadata columns to the Seurat object:

```{r metadata}
sample_meta <- sample_info[!names(sample_info) %in% 
                              c('sample',
                                'HDF5', 'RDS', 'Gene.Expression', 'Peaks', 'TF', 
                                'fragments', 'singlecell')]


# Add metadata variables based on the sample table
for (i in 1:length(sample_meta)) {
    group_name <- names(sample_meta)[i]
    print(group_name)
    map_df <- data.frame(
        unit = unlist(sample_info[['sample']]),
        group_name = unlist(sample_info[[group_name]])
    )
    group_elements <- plyr::mapvalues(
        x = seurat$orig.ident, 
        from = map_df$unit, 
        to = map_df$group_name
    )
    seurat <- AddMetaData(
        object = seurat, 
        metadata = group_elements, 
        col.name = group_name
    )
}


# map 10X barcode suffix to library_id
barcode_suffix <- as.integer(gsub('.*-([0-9]+)$', '\\1', colnames(seurat)))
if (class(barcode_suffix) != "integer") {
    stop("Barcode suffix is not integer!")
}

# Add the barcode suffix to metadata
seurat <- AddMetaData(
    object = seurat,
    metadata = barcode_suffix,
    col.name = 'bc_suffix'
)

if (params[['aggregates_tsv']] != "") {
    # add any remaining library metadata in aggrates.tsv
    cols <- c("sample", "replicate", "library_id", "bc_suffix")
    aggregates_meta <- c(
        "library_id", 
        colnames(aggregates_info)[!(colnames(aggregates_info) %in% cols)]
        )

    # if any aggregates col names in seurat, make aggregates colnames unique
#     non_unique <- colnames(seurat) %in% aggregates_meta
#     if (length(non_unique) > 1) {
#         idx <- which(aggregates_meta %in% non_unique)
#         aggregates_meta
#     }

    for (i in 1:length(aggregates_meta)) {
        group_name <- aggregates_meta[i]
        print(group_name)
        map_df <- data.frame(
            unit = unlist(aggregates_info[['bc_suffix']]),
            group_name = unlist(aggregates_info[[group_name]])
        )
        group_elements <- plyr::mapvalues(
            x = barcode_suffix, 
            from = map_df$unit, 
            to = map_df$group_name
        )
        # ensure aggregates.tsv metadata names are not duplicates of samples.tsv
        if (group_name %in% colnames(seurat@meta.data)) {
            group_name <- paste0(group_name, ".", i)
        }
        seurat <- AddMetaData(
            object = seurat, 
            metadata = group_elements, 
            col.name = group_name
        )
    }
}
print("First two rows for the metadata prepared:")
print(head(seurat@meta.data, 2))
```


## Save Seurat object

```{r save}
cat('Updated Seurat object:')
print(seurat)

saveRDS(object=seurat, file=params[['seurat']])
print(paste("Seurat obj saved:", params[['seurat']]))
```


## Session Info

```{r session_info}
sessionInfo()

```



