---
title: "Integration"
output:
    html_document:
            toc: true
            toc_float: true
            code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    warning=FALSE,
    message=FALSE
    )

```

```{r packages_and_functions, include=FALSE}
library(Signac)
library(Seurat)
library(GenomeInfoDb)
library(knitr)
library(future)
set.seed(1234)

source('common.R')
```




## Parallelization

```{r parallelize, include=FALSE}

# Assign the number of cores for parallelization
ncpus <- future::availableCores()

# Adjust max memory allowed
options(future.globals.maxSize=Inf,
        mc.cores=ncpus)

# Parallelize the run
plan("multicore", workers=ncpus)


```


## Set variables

We'll get all the sample and replicate specific variables listed in samples.tsv and units.tsv, respectively. We'll also set paths to input files.

```{r variables}
config_path <- params[['config']]
```


## Read in data

Read in all Seurat objects and combine in a list.

```{r input}
seurat_obj <- readRDS(params[['seurat_in']])
cat('Input Seurat object: \n')
print(seurat_obj)

tryCatch(expr={ DefaultAssay(seurat_obj) <- "Peaks" },
         error=function(x) { DefaultAssay(seurat_obj) <- "Gene.Expression" })

config <- yaml::read_yaml(file = config_path)

integrate_activate <- config[['integrate']][['activate']]
memory_MB <- config[['integrate']][['memory_MB']]
atac_integrate_embeddings <- config[['integrate']][['atac_integrate_embeddings']]
```


# Group data for normalization

We'll iterate over a single group listed in the 'normalize' section of the yaml config file (Snakemake will choose which assay to use right now). For every group found, we'll look for the selected assay and normalize using a user selected method.

```{r group}
target <- basename(params[['seurat_out']])
target <- gsub("^(.*).rds", "\\1", target)
print("Integrating for group:")
print(target)

groups <- config[['integrate']][['groups']]

idx <- which(names(groups) %in% target)
 
if (length(idx) == 0) {
    print("Group is not present in config's normalize groups list.")
    q()
} else if (length(idx) > 1) {
    print("You are trying to normalize group/assay multiple times. Please fix.")
    q()
}

group_name <- names(groups)[idx]
print('Current group:')
print(group_name)

group <- groups[[idx]]
```


# Integrate Data

We'll pick a metadata column in the Seurat object, split the input Seurat object based on metadata labels within the column of interest, and integrate between those groups.

```{r params_list}
split_by <- group[['split_by']]
	
ref_label <- group[['reference']][['label']]
query_label <- group[['query']][['label']]
features_use <- c(ref_label, query_label)
    
ref_assay <- group[['reference']][['assay_name']]
query_assay <- group[['query']][['assay_name']]
assays <- c(ref_assay, query_assay)
    
# new_assay_name = group[['new_slot_name']]
norm_method <- group[['norm_method']]
    
integrate_dims <- group[['integrate_dims']]



```


```{r prep_integration}
# Split by modalities (aka Seurat assays, aka 10X Genomics feature types) using Seurat metadata column 
print('Current integration group:')
print(group_name)

seurat_list <- SplitObject(
    object = seurat_obj,
    split.by = split_by
)

if (length(seurat_list) == 1) {
    print('Provided metadata column has only one group. Integration requires 2 or more.')
    print('Nothing to integrate. Quiting.')
    q()
}

print('list names before subsetting')
print(seurat_list)

# resulting list from 'SplitObject' may have more elements than want to integrate
# for each metadata column, subset list for metadata values
seurat_list <- seurat_list[features_use]

print('list names after subsetting')
print(seurat_list)
print(names(seurat_list))

print('setting default assays')
assays

seurat_list <- lapply(seq_along(seurat_list), function(x) {
    seurat_obj <- seurat_list[[x]]
    DefaultAssay(seurat_obj) <- assays[x]
    return(seurat_obj)
})

names(seurat_list) <- features_use

print('list names after setting default assays')
print(seurat_list)
```


```{r normalize}
# Normalize data
if (norm_method == 'log') {
    print("Performing log normalization.")
    
    ################
    
    # Find common features if integrating different assays
    # As with SCTransform
    
    # doesn't seem needed????
    # SelectIntegrationFeatures picks features present in both/all assays
    
    ################
    
    # normalize and identify variable features for each dataset independently
    seurat_list <- lapply(seq_along(seurat_list), function(x) {
        seurat_obj <- seurat_list[[x]]
        
        seurat_obj <- NormalizeData(
            object = seurat_obj,
            normalization.method = 'LogNormalize'
        )
        seurat_obj <- FindVariableFeatures(
            object = seurat_obj,
            selection.method = "vst",
            nfeatures = 2000
        )
        
        return(seurat_obj)
    })
    
    features <- SelectIntegrationFeatures(
        object.list = seurat_list,
        assay = assays
    )
    
} else if (norm_method == 'sct') {
    print("Performing SCTransform normalization.")
    
    # Integration step for SCTransform (and ATAC) require equal feature names (genes, peaks, etc)
    # if length of assays is 1 (i.e. assays = 'RNA'), 
    # rownames are already equal from merging Seurat objects in 'combine_seurat.R'
    if (length(unique(assays)) > 1) {
        print("More than one assay provided.")
        seurat_list <- lapply(seq_along(seurat_list), function(x) {
            seurat_obj <- seurat_list[[x]]
            
            assay <- assays[x]
            counts <- seurat_obj[[assay]]@counts
            obj <- Seurat::CreateAssayObject(counts = counts)
            seurat_obj[['PrepSCT']] <- obj
            
            return(seurat_obj)
        })
        
        print('merging list')
        if (length(seurat_list) == 1) {
            joint_obj <- seurat_list[[1]]
        } else if (length(seurat_list) == 2) {
            joint_obj <- merge(seurat_list[[1]], seurat_list[[2]])
        } else if (length(seurat_list) > 2) {
            joint_obj <- merge(seurat_list[[1]], seurat_list[2:length(seurat_list)])
        }
        
        print('splitting object')
        seurat_list <- SplitObject(
            object = joint_obj, 
            split.by = group_name
        )
        
        print('setting default assay')
        seurat_list <- lapply(seq_along(seurat_list), function(x) {
            seurat_obj <- seurat_list[[x]]
            DefaultAssay(seurat_obj) <- 'PrepSCT'
            return(seurat_obj)
        })
        
        assays = rep('PrepSCT', length(assays))
    }
    
    # sctransform::vst requires cells (columns) sum up to a value > 0
    # https://github.com/ChristophH/sctransform/issues/86
    seurat_list <- lapply(seq_along(seurat_list), function(x) {
        seurat_obj <- seurat_list[[x]]
        
        assay <- assays[x]
        cell_counts <- colSums(seurat_obj[[assay]]@counts)
        idx <- which(cell_counts == 0)
        
        seurat_obj <- subset(seurat_obj, cells = idx, invert = TRUE)
    })
    
    print('prepping SCTransform.')
    seurat_list <- lapply(seq_along(seurat_list), function(x) {
        seurat_obj <- seurat_list[[x]]
        
        assay = assays[x]
        seurat_obj <- SCTransform(
            object = seurat_obj,
            assay = assay
            )
        return(seurat_obj)
    })
    
    print('Finding integration features.')
    features <- SelectIntegrationFeatures(
        object.list = seurat_list,
        nfeatures = 3000
    )
    
    print('Preparing for SCT integration.')
    seurat_list <- PrepSCTIntegration(
        object.list = seurat_list,
        anchor.features = features
    )
    
} else if (norm_method == "lsi") {
    print("Performing LSA normalization.")
    
    if (atac_integrate_embeddings) {
        print('merging list')
        if (length(seurat_list) == 1) {
            joint_obj <- seurat_list[[1]]
        } else if (length(seurat_list) == 2) {
            joint_obj <- merge(seurat_list[[1]],
                               seurat_list[[2]])
        } else if (length(seurat_list) > 2) {
            joint_obj <- merge(seurat_list[[1]],
                               seurat_list[2:length(seurat_list)])
        }
        print('joint object')
        print(joint_obj)
        
        DefaultAssay(joint_obj) <- 'Peaks'
        
        # process the combined dataset, need LSI for full data later
        joint_obj <- FindTopFeatures(
            object = joint_obj,
            min.cutoff = 10
        )
        joint_obj <- RunTFIDF(
            object = joint_obj
        )
        joint_obj <- RunSVD(
            object = joint_obj
        )
    } else {
        seurat_list <- lapply(seq_along(seurat_list), function(x) {
            seurat_obj <- seurat_list[[x]]
            
            seurat_obj <- RunTFIDF(seurat_obj)
            
            # The low dynamic range of scATAC-seq data makes it challenging
            # to perform variable feature selection, as we do for scRNA-seq.
            # Instead, we can choose to use only the top n% of features (peaks) for dimensional reduction.
            seurat_obj <- FindTopFeatures(
                object = seurat_obj,
                min.cutoff = 10
            )
            
            return(seurat_obj)
        })
        
        features <- SelectIntegrationFeatures(
            object.list = seurat_list,
            nfeatures = floor(0.25 * nrow(seurat_list[[1]]))
        )
        
        seurat_list = lapply(seq_along(seurat_list), function(x) {
            seurat_obj <- seurat_list[[x]]
            
            # need to have dim reduction in at least one assays,
            # since specifying weight.reduction="lsi" in IntegrateData()
            seurat_obj <- RunSVD(seurat_obj)
            
            return(seurat_obj)
        })
    }
} else {
    print('Invalid choice for integration normalization.')
    print('please select log, sct or tfidf methods and re-run this step')
}


```


```{r integrate}
# https://satijalab.org/seurat/v3.0/future_vignette.html
# set global variable size to 5-35 GiBs depending on number cells integrating.
# if (!is.null(memory_MB) && memory_MB > 500) {
#     options(future.globals.maxSize = memory_MB * 1024^2)
# }

# Set the number of neighbors when weighting anchors
k.weight <- ifelse(config[['dataset_size']][['toydataset']],
                   config[['dataset_size']][['toy_k']],   # 10 by default
                   100)

if (is.null(integrate_dims)) {
    integrate_dims = 1:30
}

if (norm_method == "log") {
    print("Performing Log integration.")
    
    anchors <- FindIntegrationAnchors(
        object.list = seurat_list,
        assay = assays,
        dims = integrate_dims,
        normalization.method = "LogNormalize",
        anchor.features = features
    )
    
    # new.assay.name = new_assay_name,
    integrated <- IntegrateData(
        anchorset = anchors,
        dims = integrate_dims,
        k.weight=k.weight,
        normalization.method = "LogNormalize"
    )
    
    DefaultAssay(integrated) <- "integrated"
    
    # Log norm method needs further scaling
    integrated <- ScaleData(
        object = integrated
    )
    
    reduction_name = 'integrated_pca'
    integrated <- RunPCA(
        object = integrated,
        reduction.name = reduction_name,
    )

} else if (norm_method == "sct") {
    print("Performing SCT integration.")
    
    anchors <- FindIntegrationAnchors(
        object.list = seurat_list,
        dims = integrate_dims,
        normalization.method = "SCT",
        anchor.features = features
    )
    # new.assay.name = new_assay_name,
    integrated <- IntegrateData(
        anchorset = anchors,
        dims = integrate_dims,
        k.weight=k.weight,
        normalization.method = "SCT"
    )
    
    DefaultAssay(integrated) <- "integrated"
    
    # !!!!!!!!!! When integrating with SCT method DO NOT scale data with ScaleData() !!!!!!!!!!
    
    reduction_name = 'integrated_pca'
    integrated <- RunPCA(
        object = integrated,
        reduction.name = reduction_name,
    )
 

} else if (norm_method == "lsi") {
    print("Performing LSA integration.")
    
    if (atac_integrate_embeddings) {
        print("Integrating embeddings.")
        # integrating only dim reduc embringings is more efficient for sparse ATAC matrices
        
        anchors <- FindIntegrationAnchors(
            object.list = seurat_list,
            anchor.features = rownames(seurat_list[[1]]),
            dims = 2:max(integrate_dims)
        )
        
        # # integrated all, even if using 2:30 for FindIntegrationAnchors
        # dims_to_integrate = seq(from = 1, to = max(dims))
        
        # integrate LSI embeddings
        # reduction_name = paste0(tolower(new_assay_name), '_lsi')


        reduction_name = 'integrated_lsi'
        integrated <- IntegrateEmbeddings(
            anchorset = anchors,
            reductions = joint_obj[['lsi']],
            new.reduction.name = reduction_name,
            dims.to.integrate = integrate_dims,
            k.weight=k.weight  # 100 by default. Adjusted to 10 if input is toydataset.
        )

    } else {
        print("Integrating peaks matrix.")
        # original Seurat integration method. Transforms matrix in active assay.
        
        # use k.filter = 'NA'. https://github.com/timoast/signac/issues/94
        anchors <- FindIntegrationAnchors(
            object.list = seurat_list,
            anchor.features = features,
            dims = 2:max(integrate_dims),
            k.filter = NA
        )

        # new.assay.name = new_assay_name,
        integrated <- IntegrateData(
            anchorset = anchors,
            dims = integrate_dims,
            weight.reduction = seurat_list[[2]][['lsi']],
            k.weight=k.weight,  # 100 by default. Adjusted to 10 if input is toydataset.
            preserve.order = TRUE
        )
        
        DefaultAssay(integrated) <- "integrated"
        
        # we now have a "corrected" TF-IDF matrix, and can run LSI again on this corrected matrix
        # reduction_name = paste0(tolower(new_assay_name), '_lsi')
        reduction_name = 'integrated_lsi'
        integrated <- RunSVD(
            object = integrated,
            reduction.name = reduction_name
        )
    }
}
```

## Double check object & dimension reduction {.tabset .tabset-pills}


### Elbow Plot

```{r finish, results='asis'}
print('final seurat objects, with integrated assay and linear dim reduciton')
print(integrated)

targets = c("pca", "lsi")
idx = grep(paste(targets,collapse="|"), names(integrated@reductions))
reduction_name = names(integrated@reductions)[idx]

max_comps = ncol(integrated[[reduction_name]]@cell.embeddings)

# Seurat::Stdev() has "pca" hard coded (from IRLBA)
# calc our own standard deviations for cell embedding, if absent (i.e. LSI)
sdev = integrated[[reduction_name]]@stdev
if (length(sdev) == 0) {
    m = integrated[[reduction_name]]@cell.embeddings
    sdev = apply(m, 2, sd)
    integrated[[reduction_name]]@stdev = sdev
}

ElbowPlot(object = integrated, ndims = max_comps, reduction = reduction_name)
```


## Save Seurat object

```{r save}

saveRDS(object=integrated, file=params[['seurat_out']])
print('Done saving list of seurat objects.')
```
max_comps = ncol(integrated[[reduction_name]]@cell.embeddings)

## Session Info

```{r session_info}
sessionInfo()
```

