---
title: "Integration"
output:
    html_document:
            toc: true
            toc_float: true
            code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    warning=FALSE,
    message=FALSE
    )

```

```{r packages_and_functions, include=FALSE}
library(tidyverse)
library(Signac)
library(Seurat)
library(GenomeInfoDb)
library(knitr)
library(future)
set.seed(1234)

source('common.R')
```

# Parallelization

```{r parallelize, include=FALSE}

# Assign the number of cores for parallelization
ncpus <- future::availableCores()

# Adjust max memory allowed
options(future.globals.maxSize=Inf,
        mc.cores=ncpus)

# Parallelize the run
plan("multicore", workers=ncpus)
```

# Set variables

We'll get all the sample and replicate specific variables listed in samples.tsv and units.tsv, respectively. We'll also set paths to input files.

```{r variables}
config_path <- params[['config']]
```


# Read in data

Read in all Seurat objects and combine in a list.

```{r input}
seurat <- readRDS(params[['seurat_in']])
print('Input Seurat object:')
print(seurat)

config <- yaml::read_yaml(file = config_path)

integrate_activate <- config[['integrate']][['activate']]
atac_integrate_embeddings <- config[['integrate']][['atac_integrate_embeddings']]
```


# Group data for normalization

We'll iterate over a single group listed in the 'normalize' section of the yaml config file 
(Snakemake will choose which assay to use right now). For every group found, we'll look for the selected 
assay and normalize using a user selected method.

```{r group}

target <- basename(params[['seurat_out']])
target <- gsub("^(.*).rds", "\\1", target)

# Retrieve configuration for integration
group <- config[['integrate']][['groups']][[target]]
assay_name <- group[['assay_name']]

split_by <- config[['integrate']][['split_by']]
norm_method <- group[['norm_method']]
integrate_dims <- group[['integrate_dims']]
integrate_method <- group[['integrate_method']]

if (is.null(integrate_dims)) {
    integrate_dims <- 1:30
}

print(paste("Integrating for group:", target))
print(paste("Integrating for assay:", assay_name))
print(paste("Integrating for metadata column:", split_by))
print(paste("Integrating for normalization method:", norm_method))
print(paste("Number of dimensions used for integration:", integrate_dims[1], "to", integrate_dims[2]))
print(paste("Integration method configured:", integrate_method))

# Set the number of neighbors when weighting anchors
k.weight <- ifelse(config[['dataset_size']][['toydataset']],
                   config[['dataset_size']][['toy_k']],   # 10 by default
                   100)
```


# Integrate Data

We'll pick a metadata column in the Seurat object and integrate between those groups. The current analysis is
configured to integrate dataset by `r split_by`. 

Integration approaches are described in the following documentations:

- scRNA-seq: [Integrative analysis in Seurat v5](https://satijalab.org/seurat/articles/seurat5_integration)
- scATAC-seq: [scATAC-seq data integration](https://stuartlab.org/signac/articles/integrate_atac)

For ATAC data normalized with
[term frequency inverse document frequency (TF-IDF)](https://stuartlab.org/signac/reference/runtfidf), 
integration is performed using the `IntegrateEmbeddings` function in 
[Signac](https://stuartlab.org/signac/). The main steps of this procedure are identical to 
[IntegrateData](https://satijalab.org/seurat/reference/integratedata) with one key distinction. 
When computing the weights matrix, the distance calculations are performed in the full space of 
integrated embeddings when integrating more than two datasets, as opposed to a reduced PCA space 
which is the default behavior in IntegrateData.


```{r sanity_check}

sanity_check <- c(
    assay= assay_name %in% names(seurat),
    dimred= names(seurat@reductions)
)

sanity_check['norm_method'] <- ifelse(
    norm_method == "sct", "SCT", ifelse(
    norm_method == "log", "LogNormalize", ifelse(
    norm_method == "lsi", "LSI", NULL)
)) 

sanity_check['N_dataset'] <- length(SplitObject(seurat, split_by))

# Break if your configured assay name is not found in your current Seurat obj
if (sanity_check['assay'] != "TRUE") {
    stop(paste(assay_name, "not found in your single cell object!"))
}

# Run dimensionality reduction if missing
if (!any(sanity_check %in% c("pca", "lsi")) & sanity_check['N_dataset'] > 1) {
    print("Reduced dimensions missing in the current single cell obj.")
    print("Running dimensionality reduction...")
    seurat <- preprocess_seurat(
        seurat_obj=seurat,
        assay=assay_name,
        norm_method=norm_method
    )
}

dim_name <- names(seurat@reductions)

if (is.null(sanity_check['norm_method'])) {
    stop("Current normalization method is not available for integration!")
}
```


```{r integrate}

integrated <- NULL

# Integration runs only when you have more than one datasets
if (sanity_check['N_dataset'] > 1) {

    print(paste("Integration performed on the following counts:", norm_method))
    if (norm_method %in% c("sct", "log")) {
        # Integrate SCT or log-transformed datasets
        integrated <- if (integrate_method == "CCAIntegration") {
            IntegrateLayers(
                seurat,
                method=CCAIntegration,
                orig.reduction="pca",
                new.reduction="integrated_pca",
                normalization.method=sanity_check["norm_method"],
                dims=integrate_dims,
                k.weight=k.weight
                )
        } else if (integrate_method == "RPCAIntegration") {
            IntegrateLayers(
                seurat,
                method=RPCAIntegration,
                orig.reduction="pca",
                new.reduction="integrated_pca",
                normalization.method=sanity_check["norm_method"],
                dims=integrate_dims,
                k.weight=k.weight
                )
        } else if (integrate_method == "HarmonyIntegration") {
            IntegrateLayers(
                seurat,
                method=HarmonyIntegration,
                orig.reduction="pca",
                new.reduction="integrated_pca",
                normalization.method=sanity_check["norm_method"],
                dims=integrate_dims,
                k.weight=k.weight
                )
        } else if (integrate_method == "FastMNNIntegration") {
            IntegrateLayers(
                seurat,
                method=FastMNNIntegration,
                orig.reduction="pca",
                new.reduction="integrated_pca",
                normalization.method=sanity_check["norm_method"],
                dims=integrate_dims,
                k.weight=k.weight
                )
        } else if (integrate_method == "scVIIntegration") {
            IntegrateLayers(
                seurat,
                method=FastMNNIntegration,
                orig.reduction="pca",
                new.reduction="integrated_pca",
                normalization.method=sanity_check["norm_method"],
                dims=integrate_dims,
                k.weight=k.weight
                )
        } else {
            stop("Configure valid integrate_method in config.yaml!")
        }

    } else if (norm_method == "lsi") {

        # Convert obj into a list of subsetted obj
        seurat_list <- SplitObject(seurat, split_by)

        anchors <- FindIntegrationAnchors(
            object.list=seurat_list,
            anchor.features=rownames(seurat_list[[1]]),
            reduction="rlsi",
            dims=2:max(integrate_dims)
            )

        integrated <- IntegrateEmbeddings(
                anchorset = anchors,
                reductions = seurat[['lsi']],
                new.reduction.name = 'integrated_lsi',
                dims.to.integrate = integrate_dims,
                k.weight=k.weight  # 100 by default. Adjusted to 10 if input is toydataset.
            )

    } else {
        stop(paste(norm_method, "is unavailable for normalization in", assay_name))
    }
} else {
    stop(paste("Not enough dataset for integration. Delete", target, "from your config.yaml!"))
}

if (!is.null(integrated)) {
    print('Final seurat objects, with integrated assay and linear dim reduction:')
    print(integrated)
}
```


# Elbow Plot

```{r finish, results='asis'}

# Determine the name of reduced dimension for plotting
targets <- c("pca", "lsi")
idx <- grep(paste(targets,collapse="|"), names(integrated@reductions))
reduction_name <- names(integrated@reductions)[idx]
reduction_name <- reduction_name[str_detect(reduction_name, "integrated")]

max_comps <- ncol(integrated[[reduction_name]]@cell.embeddings)

# Seurat::Stdev() has "pca" hard coded (from IRLBA)
# calc our own standard deviations for cell embedding, if absent (i.e. LSI)
sdev <- integrated[[reduction_name]]@stdev
if (length(sdev) == 0) {
    m <- integrated[[reduction_name]]@cell.embeddings
    sdev <- apply(m, 2, sd)
    integrated[[reduction_name]]@stdev <- sdev
}

ElbowPlot(
    object=integrated,
    ndims=max_comps,
    reduction=reduction_name)
```


# Save Seurat object

```{r save}
saveRDS(object=integrated, file=params[['seurat_out']])
print(paste("Seurat obj saved in", params[['seurat_out']]))
```

## Session Info

```{r session_info}
sessionInfo()
```

