---
title: "Merge Following QC"
output:
    html_document:
            toc: true
            toc_float: true
            code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    warning=FALSE,
    message=FALSE
    )
```

```{r packages_and_functions, include=FALSE}
library(tidyverse)
library(GenomeInfoDb)
library(Signac)
library(Seurat)
library(future)

source('common.R')
```

## Design of this script 

This script is designed as a replacement of the three rules below in the snakemake pipeline:    

- `merge_macs_prep`
- `macs`
- `add_macs_peaks`


This script is executed if macs2 is unset in the pipeline. Given that, this script includes essential steps from `merge_macs_prep` and `add_macs_peaks`.

## Parallelization

```{r future_parallelization, include=TRUE}
# Assign the number of cores for parallelization
ncpus <- future::availableCores()

# Adjust max memory allowed
options(future.globals.maxSize=48000 * 1024^2,
        mc.cores=ncpus)

# Parallelize the run
plan("multicore", workers=ncpus)

```

## Set variables

We'll get all the sample and replicate specific variables listed in samples.tsv and units.tsv, respectively. We'll also set paths to input files.

```{r variables}

seurat_inputs <- unlist(strsplit(params[['seurats']], ' '))
print('Input Seurat objects:')
print(seurat_inputs)

```

## Read in data

Read in all Seurat objects and combine in a list.

```{r input}
# Import seurat objects
seurat_list <- lapply(seurat_inputs, readRDS)

print('Input sample list:')
print(seurat_list)

# Import annotation (GRanges object)
annotation <- readRDS(params[["annotation"]])
```

## Update ZINBA peak coordinates (if necessary)

- Peaks are called by ZINBA during the cellranger-arc run (saved in `atac_peaks.bed`). See more info 
about [cellranger-arc-mediated peak calling](https://support.10xgenomics.com/single-cell-multiome-atac-gex/software/pipelines/latest/output/peaks).

We will check if there are more than one sample with a Peaks assay, and if so, take the union of all loci 
and update peak x cell matrix. Unless all peaks have the same genomic coordinates, downstream processing 
fails. A better way to do handle this is to use cellranger-arc aggr, which merges peaks across samples 
and re-counts fragments which those loci.

```{r update_zinba_peaks}

# Create a logical vector indicating the presence of `Peaks`
peaks_idx <- sapply(seq_along(seurat_list), function(x) {
    any(names(seurat_list[[x]]) %in% 'Peaks')
})

# peaks present in more >1 sample? If so, update peaks coords across samples
if (sum(peaks_idx, na.rm = TRUE) > 1) {
    
    # make a list of all peaks coords
    peaks <- seurat_list[peaks_idx] %>%
        lapply(function(o) o[['Peaks']]@ranges)

    # merge all peak coords
    merged_peaks <- GenomicRanges::reduce(Reduce("c", peaks))

    print('Updated peaks for use with all Peaks X Cells matrices:')
    print(merged_peaks)
    print("-----------------------------------------------------------")

    # update seurat objects in list only if the have peaks
    seurat_list <- lapply(seq_along(seurat_list), function(x) {
        seurat <- seurat_list[[x]]

        if (peaks_idx[x]) {
            print(paste("ATAC sample no.", x))
            print('peaks before updating coords:')
            print(head(rownames(seurat[['Peaks']])))

            # Update the feature x cell matrix based on re-aligned peak coordinates
            peak_counts <- FeatureMatrix(fragments=Fragments(seurat),
                                         features=merged_peaks,
                                         cells=colnames(seurat))

            # Build a new peak obj using the updated matrix
            new_peaks <- Signac::CreateChromatinAssay(
                counts = peak_counts, 
                fragments = Fragments(seurat),
                annotation = annotation
            )

            # set to non-Peaks assay
            # even single-assay Peaks kit from 10X Genomics will 
            # have Gene.Activity and/or TF assays
            assay_idx <- !names(seurat) %in% "Peaks"

            # overwrite existing Peaks assay with single sample peaks
            seurat[['Peaks']] <- new_peaks

            print('peaks after updating coords:')
            print(head(rownames(seurat[['Peaks']])))

            print("-----------------------------------------------------------")

        }

        return(seurat)
    })
} else {
    print('Not updating ZINBA peaks. 0-1 Peaks X Cells matrices in samples.')
}
```

## Merge all individual seurat objects into joint object


```{r merge}

# Merge seurat objects
seurat <- merge(seurat_list[[1]], seurat_list[-1])

# Set default assay properly
if (any(names(seurat@assays) %in% 'Gene.Expression')) {
    DefaultAssay(seurat) <- 'Gene.Expression'
}

# for some reason, ChromatinAssay class are coerced to Assay class
if (any(names(seurat) %in% 'Peaks')) {
    # Ensure that data structure is unchanged after merge
    if (class(seurat[['Peaks']]) != "ChromatinAssay") {
        seurat[['Peaks']] <- as.ChromatinAssay(seurat[['Peaks']])
    }
    if (!identical(seurat[['Peaks']]@annotation, annotation)) {
        seurat[['Peaks']]@annotation <- annotation
    }

    # fragments slot is a list of objects of Fragments class
    # seurat[['Peaks']]@fragments consist of fragments from each ATAC sample
    for (x in seq_along(seurat_list)) {
        seurat_obj <- seurat_list[[x]]
        if (any(names(seurat_obj) %in% 'Peaks')) {
            # Ensure that the fragments slot contains fragments from each sample
            if (length(Fragments(seurat_obj)) == 1 & 
                !identical(seurat[['Peaks']]@fragments[[1]], Fragments(seurat_obj)[[1]])) {
                seurat[['Peaks']]@fragments <- lapply(
                    seurat[['Peaks']]@fragments,
                    function(x) Fragments(seurat_obj)[[1]]
                )
            } else if (length(Fragments(seurat_obj)) > 1) {
                # fix later. Below creates nested list of frag objects
                # seurat[['Peaks']]@fragments[[i]] = Fragments(seurat_obj)
                stop("More than 1 fragments are found in seurat[['Peaks']]!")
            }
        }
    }

    print("Fragments present:")
    print(seurat[['Peaks']]@fragments)

    print('Updated peaks assay')
    print(seurat[['Peaks']])
}

print('Merged, unintegrated Seurat object:')
print(seurat)
```

## Save Seurat object

```{r save}
saveRDS(seurat, params[['seurat']])
print(paste('Updated Seurat obj saved in', params[['seurat']]))
```


## Session Info

```{r session_info}
sessionInfo()
```



