---
title: "Merge Following QC"
output:
    html_document:
            keep_md: yes
            toc: true
            toc_float: true
            code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    warning=FALSE,
    message=FALSE,
    results='hide'
    )
```

```{r packages_and_functions, include=FALSE}
library(GenomeInfoDb)
library(Signac)
library(Seurat)
library(future)

source('common.R')
```

## Design of this script 

This script is designed as a replacement of the three rules below in the snakemake pipeline:    

- `merge_macs_prep`
- `macs`
- `add_macs_peaks`


This script is executed if macs2 is unset in the pipeline. Given that, this script includes essential steps from `merge_macs_prep` and `add_macs_peaks`.

## Parallelization

```{r future_parallelization, include=TRUE}
# Assign the number of cores for parallelization
ncpus <- future::availableCores()

# Adjust max memory allowed
options(future.globals.maxSize=48000 * 1024^2,
        mc.cores=ncpus)

# Parallelize the run
plan("multicore", workers=ncpus)

```




## Set variables

We'll get all the sample and replicate specific variables listed in samples.tsv and units.tsv, respectively. We'll also set paths to input files.

```{r variables}
seurat_inputs = unlist(strsplit(params[['seurats']], ' '))
cat('Input Seurat objects: \n')
print(seurat_inputs)






```


## Read in data

Read in all Seurat objects and combine in a list.

```{r input}
# Import seurat objects
seurat_list <- lapply(seq_along(seurat_inputs), function(x){
	s_in = readRDS(seurat_inputs[x])
	return(s_in)
})
cat('Input sample list: \n')
print(seurat_list)

# Import annotation (GRanges object)
annotation <- readRDS(params[["annotation"]])
```

## Update ZINBA peak coordinates (if necessary)

- Peaks are called by ZINBA during the cellranger-arc run (saved in `atac_peaks.bed`). See more info about [cellranger-arc-mediated peak calling](https://support.10xgenomics.com/single-cell-multiome-atac-gex/software/pipelines/latest/output/peaks)

We will check if there are more than one sample with a Peaks assay, and if so, take the union of all loci and update peak x cell matrix. Unless all peaks have the same genomic coordinates, downstream processing fails. A better way to do handle this is to use cellranger-arc aggr, which merges peaks across samples and re-counts fragments which those loci.

```{r update_zinba_peaks}
# any peaks present?
peaks_idx = sapply(seq_along(seurat_list), function(x) {
    idx = any(names(seurat_list[[x]]) %in% 'Peaks')
    return(idx)
})

# peaks present in more >1 sample? If so, update peaks coords across samples
if (sum(peaks_idx, na.rm = TRUE) > 1) {
    
    # make a list of all peaks coords
    peaks = lapply(seq_along(peaks_idx), function(x) {
        if (peaks_idx[x]) {
            regions = seurat_list[[x]][['Peaks']]@ranges
            return(regions)
        } else {
            regions = NA
        }
        return(NA)
    })
    
    # merge all peak coords
    idx = which(!is.na(peaks))
    peaks = peaks[idx]
    merged_peaks = GenomicRanges::reduce(Reduce("c", peaks))
    print('Updated peaks for use with all Peaks X Cells matrices:')
    print(merged_peaks)
    
    # update seurat objects in list only if the have peaks
    seurat_list = lapply(seq_along(seurat_list), function(x) {
        seurat = seurat_list[[x]]
        
        if (peaks_idx[x]) {
            print('peaks before updating coords:')
            print(head(rownames(seurat[['Peaks']])))
                
            peak_counts <- FeatureMatrix(fragments=Fragments(seurat),
                                         features=merged_peaks,
                                         cells=colnames(seurat))
            
            new_peaks <- Signac::CreateChromatinAssay(
                counts = peak_counts, 
                fragments = Fragments(seurat),
                annotation = annotation
            )
            
            # set to non-Peaks assay
            # even single-assay Peaks kit from 10X Genomics will 
            # have Gene.Activity and/or TF assays
            assay_names = names(seurat)
            assay_idx = !names(seurat) %in% "Peaks"
            
            # overwrite existing Peaks assay with single sample peaks
            seurat[['Peaks']] = new_peaks
            
            # if (length(assay_names[assay_idx]) > 0) {
            #     assay_use = assay_names[assay_idx][1]
            #     DefaultAssay(seurat) = assay_use
            #     
            #     seurat[['Peaks']] = NULL
            #     seurat$nCount_Peaks = NULL
            #     seurat$nFeature_Peaks = NULL
            #     seurat[['Peaks']] = new_peaks
            # } else {
            #     seurat[['Peaks']] = new_peaks
            # }
            
            print('peaks after updating coords:')
            print(head(rownames(seurat[['Peaks']])))
            print(head(seurat@meta.data))
        }
        
        return(seurat)
    })
} else {
    print('Not updating ZINBA peaks. 0-1 Peaks X Cells matrices in samples.')
}
```



## Merge all individual seurat objects into joint object


```{r merge}
if (length(seurat_list) == 1) {
	seurat = seurat_list[[1]]
} else if (length(seurat_list) == 2) {
	seurat = merge(seurat_list[[1]], seurat_list[[2]])
} else if (length(seurat_list) > 2) {
	seurat = merge(seurat_list[[1]], seurat_list[2:length(seurat_list)])
}

if (any(names(seurat@assays) %in% 'Gene.Expression')) {
	DefaultAssay(seurat) = 'Gene.Expression'
}

# for some reason, ChromatinAssay class are coerced to Assay class
if (any(names(seurat) %in% 'Peaks')) {
	seurat[['Peaks']] = as.ChromatinAssay(seurat[['Peaks']], annotation = annotation)
	
	# seurat[['Peaks']]@annotation <- params[["annotation"]]
	
	# fragments slot is a list of objects of Fragments class
	i = 1
	for (x in seq_along(seurat_list)) {
	    seurat_obj = seurat_list[[x]]
	    if (any(names(seurat_obj) %in% 'Peaks')) {
	        cat('\nPeaks found in:', x, '\n')
	        if (length(Fragments(seurat_obj)) == 1) {
	            seurat[['Peaks']]@fragments[[i]] = Fragments(seurat_obj)[[1]]
	        } else if (length(Fragments(seurat_obj)) > 1) {
	            # fix later. Below creates nested list of frag objects
	            # seurat[['Peaks']]@fragments[[i]] = Fragments(seurat_obj)
	        }
	        i = i + 1
	    }
	}
	
	seurat[['Peaks']]@fragments

	print('updated peaks assay')
	print(seurat[['Peaks']])
	# print(seurat[['Peaks']]@fragments[[1]]@path)
	# print(seurat[['Peaks']]@fragments[[1]]@hash)
	# print(head(seurat[['Peaks']]@fragments[[1]]@cells))
}

cat('Merged, unintegrated Seurat object: \n')
print(seurat)
```



## Save Seurat object

```{r save}
print('Updated Seurat object:')
print(seurat)

saveRDS(seurat, params[['seurat']])
print('Done saving seurat object.')
```


## Session Info

```{r session_info}
sessionInfo()
```



