---
title: "Normalization and Dimentionality Reduction"
output:
    html_document:
            toc: true
            toc_float: true
            code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    warning=FALSE,
    message=FALSE
    )

```

```{r packages_and_functions, include=FALSE}
library(Signac)
library(Seurat)
library(GenomeInfoDb)
library(knitr)
library(future)
set.seed(1234)

source('common.R')
```

## Parallelization

```{r parallelize, include=FALSE}

# Assign the number of cores for parallelization
ncpus <- future::availableCores()

# Adjust max memory allowed
options(future.globals.maxSize=40000 * 1024^2,
        mc.cores=ncpus)

# Parallelize the run
plan("multicore", workers=ncpus)


```


## Set variables

We'll get all the sample and replicate specific variables listed in samples.tsv and units.tsv, respectively. We'll also set paths to input files.

```{r variables}
config_path <- params[['config']]
```


## Read in data

Read in all Seurat objects and combine in a list.

```{r input}

seurat <- readRDS(params[['seurat_in']])
print('Input Seurat object:')
print(seurat)


tryCatch(expr={ DefaultAssay(seurat) <- "Peaks" },
         error=function(e) { DefaultAssay(seurat) <- "Gene.Expression" }
)


config <- yaml::read_yaml(file = config_path)
```


## Group data for normalization

We'll iterate over a single group listed in the 'normalize' section of the yaml config file (Snakemake will choose which assay to use right now). For every group found, we'll look for the selected assay and normalize using a user selected method.



```{r group}
target <- basename(params[['seurat_out']])
target <- gsub("^(.*).rds", "\\1", target)
print(paste("Normalizing for group:", target))

split_by <- config[['normalize']][['split_by']]

groups <- config[['normalize']][['groups']]

idx <- which(names(groups) %in% target)
 
if (length(idx) == 0) {
    stop("Group is not present in config's normalize groups list.")
} else if (length(idx) > 1) {
    stop("You are trying to normalize group/assay multiple times. Please fix.")
}

group_name <- names(groups)[idx]
print(paste('Current group:', group_name))

group <- groups[[idx]]

assay <- group[['assay_name']]
norm_method <- group[['norm_method']]

print(paste('Current assay:', assay))
print(paste('Current norm_method:', norm_method))
```


## Check if assay is present in data

We'll also look for cells that have zero or NA values in the metadata column for the selected group assay. 
This is needed because when merging Seurat objects (necessary for multiple samples/replicates), samples 
that have zero counts for an assay (Peaks column for GEX samples) are assigned NA in the respective metadata 
column. This breaks most downstream processes.

```{r check}
# remove cells with 0 counts for the selected assay
var_name <- paste0('nCount_', assay)
print(paste("Cells with zero or missing values are removed based on", var_name))

# Create a logical vector for each cell having non-zero counts
new_idx <- tryCatch(expr={
             fetch <- FetchData(object = seurat, vars = var_name)
             !is.na(fetch) & (fetch > 0) },
        error=function(e) { idx }
        )
```


## Normalize and Linear Dimension Reduction

Now we'll iterate over the list and perform normalization for each assay. For each element in the list of 
seurat objects (grouped by assay), we'll call `common.R::preprocess_seurat()` which performs log, clr, 
or lsi normalization. For each assay a unique dimensional reduction name is created, in case we need to have 
multiple reductions (umap for GEX and umap for ATAC, etc)

```{r normalize}

if (length(new_idx) > 0 & !identical(idx, new_idx)) {
    # Subset cells with non-zero counts if present
    seurat_sub <- seurat[,which(x = new_idx)]

    # Ensure to set default assay properly
    DefaultAssay(seurat_sub) <- assay

    # Run dimensionality reduction based on configuration
    seurat_sub <- preprocess_seurat(
        seurat_obj = seurat_sub,
        assay = assay,
        norm_method = norm_method)
} else {
    seurat_sub <- seurat
    print('Assay not present in dataset. Skipping.')
}

print('Reduced dimensions added:')
print(seurat_sub@reductions)
```


## Double check object & dimension reduction

### Elbow Plot

```{r final}
print('Final seurat objects, with normalized assay and linear dim reduciton:')
print(seurat_sub)

# Find the primary dim reduction approach
targets <- c("pca", "lsi")
idx <- grep(paste(targets,collapse="|"), names(seurat_sub@reductions))
reduction_name <- names(seurat_sub@reductions)[idx]

if (!is.null(reduction_name)) {
    # Calculate the number of dimensions calculated
    max_comps <- ncol(seurat_sub[[reduction_name]]@cell.embeddings)
    # Create an elbow plot
    ElbowPlot(object = seurat_sub, ndims = max_comps, reduction = reduction_name)
}

```


## Save Seurat object

```{r save}
print(paste("Updated Seurat object saved in", params[["seurat_out"]]))
saveRDS(object=seurat_sub, file=params[['seurat_out']])
```


## Session Info

```{r session_info}
sessionInfo()
```

