---
title: "QC"
output:
    html_document:
        toc: true
        toc_float: true
        code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    warning=FALSE,
    message=FALSE
    )

```

```{r packages_and_functions, include=FALSE}
library(tidyverse)
library(ggplot2)
library(Signac)
library(Seurat)
library(future)

source('common.R')
```

## Parallelization

```{r future_parallelization, include=FALSE}
# Assign the number of cores for parallelization
ncpus <- future::availableCores()

# Adjust max memory allowed
options(future.globals.maxSize=48000 * 1024^2,
        mc.cores=ncpus)

# Parallelize the run
plan("multicore", workers=ncpus)

```

## Set variables

We'll get all the sample and replicate specific variables listed in samples.tsv and units.tsv, respectively. We'll also set paths to input files.

```{r variables}
config_path <- params[['config']]
```


## Read in data

Read in all Seurat objects and combine in a list.

```{r input}
# Import input Seurat obj
seurat <- readRDS(params[['seurat_in']])
config <- yaml::read_yaml(file = config_path)

print(paste("Input Seurat obj path:", params[['seurat_in']]))

print("Imported Seurat obj:")
print(seurat)
```


## Impute metadata

```{r impute_metadata}

# Impute NA values
# (This step is not needed for full datasets)
print(paste("Number of missing values in metadata:", sum(is.na(seurat@meta.data))))
seurat@meta.data[is.na(seurat@meta.data)] <- 0  # Impute

print("Missing values have been imputed to zero.")
```


## Quality Control

We'll iterate over features listed in the Seurat metadata columns and remove cells with counts below lower 
limits and/or above a high threshold. There are 2 ways to select lower and/or upper limits. The default 
behavior is to get summary statistics for each feature and then select cells with counts more than X 
standard deviations from the mean (default is +/-3 sd). The function also supports using Tukey's fence 
method based on quartiles. Alternatively, we can provide lower and or upper values (integer or double) 
and a hard cutoff will be applied.

For example, we can select cells with Gene.Expression n_Features outside of 3 sd from the mean, Peaks 
n_Counts greater than 1000, and cells with mitochondrial % less than 10% (progenitors tend to be mitoticlly 
active and have a higher % mito genes than many differentiated cells).


```{r qc_metrics}

# Retrieve QC configuration
remove_outliers <- config[['qc']][['remove_outliers']]
outlier_method <- config[['qc']][['rm_outliers_method']]
meta_labels <- config[['qc']][['meta_labels']]
lower <- unlist(config[['qc']][['lower']])
upper <- unlist(config[['qc']][['upper']])

print(paste("Remove outliers?:", remove_outliers))
print(paste("Removal method:", outlier_method))
print(paste("Configured QC metrics:", paste(meta_labels, collapse=", ")))
print("Lower limits:")
print(lower)
print("Upper limits:")
print(upper)

# are use selected metadata labels present for filtering?
if (is.null(meta_labels)) {
    meta_labels <- c('nFeature_Gene.Expression', 'nCount_Gene.Expression', 'percent.mt')
}

if (any(names(seurat@meta.data) %in% meta_labels)) {
    features_present <- TRUE
    idx <- names(seurat@meta.data) %in% meta_labels
    meta_labels <- names(seurat@meta.data)[idx]
    print(paste("QC metrics present in the current obj:", paste(meta_labels, collapse=", ")))
} else {
    print('Configured QC metrics are not found in the current obj.')
    features_present <- FALSE
}

```

```{r plot_qc}
Vln.list <- list()
if (features_present) {
    # Create exploratory violin plot before filtering
    Vln.list[['before_filtering']] <- VlnPlot(
        object = seurat,
        features = meta_labels,
        ncol = length(meta_labels),
        pt.size = 0,
        group.by = 'orig.ident'
    ) & theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        plot.title = element_text(size=12))


    if (remove_outliers) {

        # get list of lower and/or upper values for each meta.data label
        min_max_list <- lapply(seq_along(unique(meta_labels)), function(x) {
            meta_col <- meta_labels[x]

            # Create a vector storing values for each metric
            counts <- unlist(seurat[[meta_col]])

            # Retrieve upper and lower thresholds
            if (any(names(lower) %in% meta_col)) {
                col_lower <- lower[names(lower) %in% meta_col]
            } else {
                col_lower <- NULL
            }

            if (any(names(upper) %in% meta_col)) {
                col_upper <- upper[names(upper) %in% meta_col]
            } else {
                col_upper <- NULL
            }

            ol <- if (all(unique(counts) == 0)) {
                print(paste('Counts vector all zeros. Returning NA in', meta_col))
                print("--------------------------------------------------------------")
                NA
            } else {
                find_outliers(
                counts = counts,
                method = outlier_method,
                k = 2.2,
                sigma = 3,
                lower = col_lower,
                upper = col_upper,
                na_rm = TRUE
                )
            }

            # Return a vector storing min and max values
            return(ol)
        }) %>%
        set_names(unique(meta_labels))

        # Remove elements returned NA
        min_max_list <- min_max_list[!is.na(min_max_list)]

        min_list <- lapply(min_max_list, `[[`, 'min') %>% set_names(names(min_max_list))
        max_list <- lapply(min_max_list, `[[`, 'max') %>% set_names(names(min_max_list))

        print("Cells are filtered based on the following lower limits:")
        for (i in 1:length(min_list)) {
            var_name <- names(min_list[i])
            var_val <- min_list[[i]]
            print(paste(var_val, "in", var_name))

            # Much like R's subset function, subset.Seurat is designed for interactive use only
            # https://github.com/satijalab/seurat/issues/1053
            expr <- FetchData(object = seurat, vars = var_name)
            seurat <- seurat[, which(x = expr > var_val)]
        }
        rm(i)
        print("--------------------------------------------------------------")


        print("Cells are filtered based on the following upper limits:")

        for (i in 1:length(max_list)) {
            var_name <- names(max_list[i])
            var_val <- max_list[[i]]
            print(paste(var_val, "in", var_name))

            expr <- FetchData(object = seurat, vars = var_name)
            seurat <- seurat[, which(x = expr < var_val)]
        }

        print("--------------------------------------------------------------")
    }

    # visualize after subsetting
    Vln.list[['after_filtering']] <- VlnPlot(
        object = seurat,
        features = meta_labels,
        ncol = length(meta_labels),
        pt.size = 0,
        group.by = 'orig.ident'
    ) & theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        plot.title = element_text(size=12))
    print('Seurat obj after QC filtering:')
    print(seurat)

} else {
    print('Seurat object without QC filtering:')
    print(seurat)
}


```



## QC plot - before filtering


```{r print_plot_I, results='asis'}

print(Vln.list[['before_filtering']])

```

## QC plot - after filtering


```{r print_plot_II, results='asis'}

print(Vln.list[['after_filtering']])

```

## Save Seurat object

```{r save}

print(paste("Seurat obj saved:", params[['seurat_out']]))
saveRDS(object=seurat, file=params[['seurat_out']])
```

## Session Info

```{r session_info}
sessionInfo()
```

