---
title: "QC"
output:
    html_document:
        toc: true
        toc_float: true
        code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    warning=FALSE,
    message=FALSE
    )

```

```{r packages_and_functions, include=FALSE}
library(ggplot2)
library(Signac)
library(Seurat)
library(future)

source('common.R')
```

## Parallelization

```{r future_parallelization, include=FALSE}
# Assign the number of cores for parallelization
ncpus <- future::availableCores()

# Adjust max memory allowed
options(future.globals.maxSize=48000 * 1024^2,
        mc.cores=ncpus)

# Parallelize the run
plan("multicore", workers=ncpus)

```

```{r params, eval=FALSE}
params = list(
    seurat_in = "results/create_seurat/toydataset-nan.seurat_raw.rds",
    config = "config/multiome-config/config.yaml",
    seurat_out = "results/qc/toydataset-nan.seurat_qc.rds",
    scriptdir = "scripts"
    )

```


## Set variables

We'll get all the sample and replicate specific variables listed in samples.tsv and units.tsv, respectively. We'll also set paths to input files.

```{r variables}
config_path = params[['config']]
```


## Read in data

Read in all Seurat objects and combine in a list.

```{r input}
seurat = readRDS(params[['seurat_in']])
cat('Input Seurat object: \n')
print(seurat)

config = yaml::read_yaml(file = config_path)
```


## Impute metadata

```{r impute_metadata}

# Impute NA values (added by Mira)
# (This step is not needed for full datasets)
sum(is.na(seurat@meta.data)) # the number of NAs
seurat@meta.data[is.na(seurat@meta.data)] <- 0  # Impute

```


## Quality Control

We'll iterate over features listed in the Seurat metadata columns and remove cells with counts below lower limits and/or above a high threshold. There are 2 ways to select lower and/or upper limits. The default behavior is to get summary statistics for each feature and then select cells with counts more than X standard deviations from the mean (default is +/-3 sd). The function also supports using Tukey's fence method based on quartiles. Alternatively, we can provide lower and or upper values (integer or double) and a hard cutoff will be applied.

For example, we can select cells with Gene.Expression n_Features outside of 3 sd from the mean, Peaks n_Counts greater than 1000, and cells with mitochondrial % less than 10% (progenitors tend to be mitoticlly active and have a higher % mito genes than many differentiated cells).


```{r qc}

remove_outliers = config[['qc']][['remove_outliers']]
outlier_method = config[['qc']][['rm_outliers_method']]
meta_labels = config[['qc']][['meta_labels']]
lower = unlist(config[['qc']][['lower']])
upper = unlist(config[['qc']][['upper']])

print(remove_outliers)
print(outlier_method)
print(meta_labels)
print(lower)
print(upper)

# are use selected metadata labels present for filtering?
if (is.null(meta_labels)) {
	meta_labels = c('nFeature_Gene.Expression', 'nCount_Gene.Expression', 'percent.mt')
}

if (any(names(seurat@meta.data) %in% meta_labels)) {
	print('Selected meta_labels from config file present in seurat object.')

	features_present = TRUE
	idx = names(seurat@meta.data) %in% meta_labels
	meta_labels = names(seurat@meta.data)[idx]
	print(meta_labels)
} else {
	print('Selected meta_labels from config file not present in seurat object.')

	features_present = FALSE
}


######
# vis and filter
######

Vln.list <- list()
if (features_present) {
	# visualize disributions before filter
	Vln.list[['before_filtering']] <- VlnPlot(
		object = seurat,
		features = meta_labels,
		ncol = length(meta_labels),
		pt.size = 0,
		group.by = 'orig.ident'
	) & theme(axis.title.x=element_blank(),
	          axis.text.x=element_blank(),
	          axis.ticks.x=element_blank())


	if (remove_outliers) {
		print("removing outliers")

		# get list of lower and/or upper values for each meta.data label
		min_max_list = lapply(seq_along(unique(meta_labels)), function(x) {
			meta_col = meta_labels[x]
			
			# print(meta_col)
			counts = unlist(seurat[[meta_col]])
			# print(head(counts))
			# print(hist(counts, breaks = 100))
			
			if (any(names(lower) %in% meta_col)) {
			  col_lower = lower[names(lower) %in% meta_col]
			} else {
			  col_lower = NULL
			}
			
			if (any(names(upper) %in% meta_col)) {
			  col_upper = upper[names(upper) %in% meta_col]
			} else {
			  col_upper = NULL
			}
			
			ol = find_outliers(
				counts = counts,
				method = outlier_method,
				k = 2.2,
				sigma = 3,
				lower = col_lower,
				upper = col_upper,
				na_rm = TRUE
			)
			return(ol)
		})
		idx = !is.na(min_max_list)
		min_max_list = min_max_list[idx]
		names(min_max_list) = meta_labels[idx]
		
		min_list = lapply(min_max_list, `[[`, 'min')
		names(min_list) = names(min_max_list)
		max_list = lapply(min_max_list, `[[`, 'max')
		names(max_list) = names(min_max_list)
		
		for (i in 1:length(min_list)) {
			var_name = names(min_list[i])
			print(var_name)
			var_val = min_list[[i]]
			print(var_val)
			
			# Much like R's subset function, subset.Seurat is designed for interactive use only
			# https://github.com/satijalab/seurat/issues/1053
			expr <- FetchData(object = seurat, vars = var_name)
			seurat = seurat[, which(x = expr > var_val)] # modified by Mira
		}
		rm(i)
		
		for (i in 1:length(max_list)) {
			var_name = names(max_list[i])
			print(var_name)
			var_val = max_list[[i]]
			print(var_val)

			expr <- FetchData(object = seurat, vars = var_name)   # modified by Mira
			seurat = seurat[, which(x = expr < var_val)]  # modified by Mira
		}
	}

	# visualize after subsetting
	Vln.list[['after_filtering']] <- VlnPlot(
		object = seurat,   # modified by Mira
		features = meta_labels,
		ncol = length(meta_labels),
		pt.size = 0,
		group.by = 'orig.ident'
	) & theme(axis.title.x=element_blank(),
	          axis.text.x=element_blank(),
	          axis.ticks.x=element_blank())

	# ggsave(
	# 	filename = snakemake@output[['vln_post_qc']],
	# 	width = 4 * length(meta_labels),
	# 	height = 6,
	# 	units = "in"
	# )

	
	print('Done filtering seurat object for each cellranger output.')
	print(seurat)

} else {
	print('Seurat object without filtering.')
	print(seurat)
}


```

## QC plot - before filtering


```{r print_plot_I, results='asis'}

print(Vln.list[['before_filtering']])

```

## QC plot - after filtering


```{r print_plot_II, results='asis'}

print(Vln.list[['after_filtering']])

```

## Save Seurat object

```{r save}
saveRDS(object=seurat, file=params[['seurat_out']])
print('Done saving seurat object.')
```


## Session Info

```{r session_info}
sessionInfo()
```

